---
title: "Detailed Insights: `r chosen_geog`"
output:
  pdf_document:
     toc: true
bibliography: "`r here::here('app/www/reports', bib)`"
engine: knitr
---

---
nocite: '@*'
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(kableExtra)
library(leaflet)

### This script writes the country insights pdf for the case of countries with AVAILABLE Tier II DATA 

#function to select only observations with chosen country 
selectGeography <- function(df){
  df %>% filter(territory == chosen_geog)
}

#country lookup table 
country_lookup <- read_csv("country_lookup.csv")

country <- country_lookup %>% 
  filter(territory == chosen_geog) %>% 
  select(country)
chosen_country <- as.character(country)

#function if needed to select only chosen country (CCA cores)
selectCountry <- function(df){
  df %>% filter(country == chosen_country)
}


#pull in curated app data, filtered to chosen territory 
app_data <- readRDS("app_data.rds") 
allstocks <- app_data$main_table %>% selectGeography()
totalstocks <- app_data$territory_tec %>% selectGeography()
map_input <- app_data$map_input #no territory listed 
map <- app_data$map_polys %>% selectGeography()


rm(app_data) #keeps the rds out of the environment 
```

```{r habitat areas, echo=FALSE}
#Area totals for individual habitats 

#mangrove values 
area_ha_mangrove <- allstocks %>% filter(habitat == "mangrove") %>% select(area_ha) 
area_low_mangrove <- allstocks %>% filter(habitat == "mangrove") %>% select(area_ha_lowerCI)
area_high_mangrove <- allstocks %>% filter(habitat == "mangrove") %>% select(area_ha_upperCI)

#marsh values
area_ha_marsh <- allstocks %>% filter(habitat == "marsh") %>% select(area_ha)
area_low_marsh <- allstocks %>%  filter(habitat == "marsh") %>% select(area_ha_lowerCI)
area_high_marsh <- allstocks %>% filter(habitat == "marsh") %>% select(area_ha_upperCI)

#seagrass values 
area_ha_seagrass <- allstocks %>% filter(habitat == "seagrass") %>% select(area_ha)
area_low_seagrass <- allstocks %>% filter(habitat == "seagrass") %>% select(area_ha_lowerCI)
area_high_seagrass <- allstocks %>% filter(habitat == "seagrass") %>% select(area_ha_upperCI)
```

```{r total stocks, echo=FALSE}
#total stocks
 countrystock <- totalstocks %>% selectGeography()
  #selectCountry()

#total stocks// per total territory
stock_country <- countrystock$Total_Stocks
stock_country_low <- countrystock$Total_Stockers_LowerCI
stock_country_high <- countrystock$Total_Stockers_UpperCI
stock_country_se <- countrystock$Total_SE

#total stocks per habitat per territory 
#this is now taken from different table 
mangrove_stocks <- allstocks %>% filter(habitat == "mangrove")
marsh_stocks <- allstocks %>% filter(habitat == "marsh")
seagrass_stocks <- allstocks %>% filter(habitat == "seagrass")

#mangrove
totalstocks_mangrove <- mangrove_stocks$Total_Stocks
totalstocks_mangrove_upper <- mangrove_stocks$area_ha_upperCI
totalstocks_mangrove_lower <- mangrove_stocks$area_ha_lowerCI
  
#marsh
totalstocks_marsh <- marsh_stocks$Total_Stocks
totalstocks_marsh_upper <- marsh_stocks$area_ha_upperCI
totalstocks_marsh_lower <- marsh_stocks$Total_Stocks_LowerCI
  
#seagrass 
totalstocks_seagrass <- seagrass_stocks$Total_Stocks
totalstocks_seagrass_upper <- seagrass_stocks$area_ha_upperCI
totalstocks_seagrass_lower <- seagrass_stocks$Total_Stocks_LowerCI

```

## Welcome!

You have reached the CCN Inventory Tool Detailed Insights Report.

Congratulations! This geography has available data.

Potential data availability within the Inventory Tool ranges from Tier I, Tier II, or Tier III Carbon stock estimates.

This document includes country-specific insights and more detailed analysis, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass habitats. This report details information for the selected geography, **`r chosen_geog`**.

Explore the rest of the dashboard for more exciting visualizations, map features and data.

Resources referenced to calculate estimates for **`r chosen_geog`** are listed below under 'References' at the end of this document.

If you have any questions or data you would like to add to the Coastal Carbon Network, please reach out to us at [CoastalCarbon\@si.edu](mailto:CoastalCarbon@si.edu){.email}.

```{r, echo=FALSE, warning=FALSE}
# "map" is created in set-up chunk using imported polygons 
# tier II, CCA samples overlay 

chosen_map_input <- map_input %>% selectGeography()

leaflet() %>%
  # basemap options
  addTiles(group = "OSM (default)") %>% 

  #add polygon layer for selected territory border
  addPolygons(data = map, weight = 2) %>% 
   # add data points (global cca samples)
        addCircleMarkers(data = chosen_map_input,
                         lng = ~longitude, lat = ~latitude, radius = 2,
                         label = ~paste(habitat, carbon_pool, "data", sep = " "),
                         group = "Samples", 
                         # eventually have veg be plotted separately for color coding
                         color = "green") 


```
This map represents the area of **`r chosen_geog`** and the locations of both soil cores and vegetation samples included in the Coastal Carbon Atlas. 


## Total Carbon Stock Estimates

Total Carbon stock estimates were calculated for each geography in total, and for mangrove, tidal marsh, and seagrass habitats when available.

We estimate that **`r chosen_geog`** contains a mean estimate of `r signif(stock_country, 4)` metric tonnes soil Carbon.

We estimate that **`r chosen_geog`** contains between `r signif(stock_country_low, 4)` to `r signif(stock_country_high, 4)` metric tonnes of soil C to a depth of 1 m, with a mean estimate of N metric tonnes C.

```{r table1 territory total, echo=FALSE}

totals <- totalstocks %>%
  selectGeography() %>%
  mutate(Country = country,
         Territory = territory,
         `Total Stocks` = signif(Total_Stocks, 4),
         `Upper CI` = signif(Total_Stockers_UpperCI, 4),
         `Lower CI` = signif(Total_Stockers_LowerCI, 4)) %>%
  select(Country, Territory, `Total Stocks`, `Upper CI`, `Lower CI`)

kable(totals, caption = "Total Geography Level Carbon Stocks")
```

This total estimate includes total mangrove soil carbon stocks,with a mean estimate of `r signif(totalstocks_mangrove, 4)` metric tonnes of soil C to a depth of 1 meter.

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
#decide if mangrove data 

  if (!is.na(totalstocks_mangrove))
     {show_text = TRUE
  } else {show_text = FALSE}
#only print table if mangrove total stock is present 

```

```{r table2 mangrove, eval=show_text, echo=FALSE, results='asis'}
mangrove_totals <- allstocks %>% filter(habitat == "mangrove") %>% 
     mutate(Country = country,
            Territory = territory,
            Habitat = habitat,
            Habitat = "Mangrove",
            `Total Stock` = signif(Total_Stocks, 4),
            `Lower CI` = signif(Total_Stocks_LowerCI, 4),
            `Upper CI` = signif(Total_Stocks_UpperCI, 4),
            `Standard Error` = signif(Total_Stocks_se, 4)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(Country, Territory, Habitat, `Total Stock`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(mangrove_totals)
```

This total estimate also includes total tidal marsh carbon stocks, ranging from `r signif(totalstocks_marsh_lower, 4)` to `r signif(totalstocks_marsh_upper, 4)` metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r signif(totalstocks_marsh, 4)`

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
#decide if marsh data 

  if (!is.na(totalstocks_marsh))
     {show_text = TRUE
  } else {show_text = FALSE}
#only print table if mangrove total stock is present 

```


```{r table3 marsh, eval=show_text, echo=FALSE, results='asis'}

marsh_totals <- allstocks %>% filter(habitat == "marsh") %>% 
     mutate(Country = country,
            Territory = territory,
            Habitat = habitat,
            Habitat = "Marsh",
            `Total Stock` = signif(Total_Stocks, 4),
            `Lower CI` = signif(Total_Stocks_LowerCI, 4),
            `Upper CI` = signif(Total_Stocks_UpperCI, 4),
            `Standard Error` = signif(Total_Stocks_se, 4)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(Country, Territory, Habitat, `Total Stock`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(marsh_totals)
```

We estimate seagrass soil carbon stocks for **`r chosen_geog`** to have a mean estimate of `r signif(totalstocks_seagrass, 4)`

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
#decide if seagrass data 

  if (!is.na(totalstocks_seagrass))
     {show_text = TRUE
  } else {show_text = FALSE}
#only print table if mangrove total stock is present 

```

```{r table4 seagrass, echo=FALSE}

seagrass_totals <- allstocks %>% filter(habitat == "seagrass") %>%
     mutate(Territory = territory,
            Habitat = habitat,
            Habitat = "Seagrass",
            `Total Stock` = signif(Total_Stocks, 4)) %>% 
  #no CI columns or se for seagrass total stocks 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data
  select(Territory, Habitat, `Total Stock`)


kable(seagrass_totals)
```

## Wetland Areas and Activities

We estimate mangrove area in **`r chosen_geog`** to be `r signif(area_low_mangrove, 4)` to `r signif(area_high_mangrove, 4)` hectares, with a mean estimate of `r signif(area_ha_mangrove, 4)` hectares according to Global Mangrove Watch @Bunting_et_al_2018.

We estimate tidal marsh area in **`r chosen_geog`** to be `r signif(area_low_marsh, 4)` to `r signif(area_high_marsh, 4)` hectares, with a mean estimate of hectares according to @Worthington_et_al_2024.

We estimate seagrass area to be **`r chosen_geog`** to be a mean of `r signif(area_ha_seagrass, 4)` hectares, according to @McKenzie_et_al_2020, which aggregates global seagrass data from a number of sources.

<!-- @McKenzie_et_al_2020 classifies seagrass area estimates as either high or medium-to-low confidence. seagrass_area_high_confidence % of the estimated seagrass area of **`r chosen_geog`** is considered high to medium confidence, while seagrass_area_low_confidence % of the estimated seagrass area is categorized as low confidence. -->

## Calculated Stocks and Emissions Factors

```{r N_cores, echo = FALSE}
#cores, sites, habitats for soil cores included in tier II calcs 
#get country 
geographies_list <- allstocks %>% select(country, territory) %>% distinct()
chosen_country <- geographies_list$country

cores <- chosen_map_input %>% filter(carbon_pool == "soil") %>% count() 
veg_samples <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% count()
sites <- chosen_map_input %>% filter(carbon_pool == "soil") %>% select(site_id) %>% distinct() %>% count()
veg_sites <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% select(site_id) %>% distinct() %>% count()
habitats <- chosen_map_input %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct() %>% count()
veg_habitat <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% select(habitat) %>% distinct() %>% count()

habitat.list <- selectCountry(map_input) %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct()
veg.habitat.list <- selectCountry(map_input) %>% filter(carbon_pool == "vegetation") %>% select(habitat) %>% distinct()
habitat.list <- as.character(habitat.list)
veg.habitat.list <- as.character(veg.habitat.list)
```

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_geog`**.

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas.

Soil core data from **`r chosen_geog`** included in the Coastal Carbon Atlas consists of `r cores` soil profiles from `r sites` sites. This data comes from `r habitats` distinct habitat type(s).

Vegetation data from **`r chosen_geog`** included in the Coastal Carbon Atlas consists of `r veg_samples` from `r veg_sites`. The vegetation data from `r chosen_geog` is sourced from `r veg_habitat`, `r veg.habitat.list`


In the case that there is not yet any country specific information in the Coastal Carbon Atlas, we instead applied IPCC Tier I estimate, which are estimated global values. IPCC Global estimates for mangrove, marsh, and seagrass ecosystems are listed below.

The table titled "Availiability of Territory-Level Data" in this section also details whether the calculated Tier II, or territory-level sediment carbon stock value is significantly different from the estimated global values. This is observed in the "Reporting Insight" column, and can be helpful in observing when a habitat has a higher or lower estimated sediment carbon stock value than the IPCC Global estimates. 

```{r IPCC, echo =FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- as_tibble(data.frame(Habitat = c("mangrove", "marsh", "seagrass"),
                                      Mean = c(386, 255, 108),
                                      `Lower_CI` = c(351,254,84),
                                      `Upper_CI` = c(424,297,139)))

kable(IPCC_tier_I, caption = "IPCC Global Value Estimates")
```

## Data Availiability by Tier - Reporting Insight

```{r tier availability, echo = FALSE}
#create table of available tier estimates and any overlap 

tier_mangrove <- allstocks %>% filter(habitat == "mangrove") %>% select(TierIorII)
tier_marsh <- allstocks %>% filter(habitat == "marsh") %>% select(TierIorII)
tier_seagrass <- allstocks %>% filter(habitat == "seagrass") %>% select(TierIorII)

tiertable <- allstocks %>% select(country, territory, habitat, TierIorII, tier_II_overlaps_TierI) %>% 
  rename(Country = country, 
         Territory = territory,
         Habitat = habitat,
         Tier = TierIorII,
         `Reporting Insight` = tier_II_overlaps_TierI) %>% 
  mutate(`Reporting Insight` = if_else(is.na(`Reporting Insight`), 
        "No Tier II data available at this time", `Reporting Insight`),
        Tier = if_else(Tier == "Tier II", "Territory-Level", "Global"))

kable(tiertable, caption = "Availiability of Territory-Level Data")

```

### Tier I Carbon Stocks - Global Estimates 

This section includes estimated global sediment carbon stocks included for **`r chosen_geog`**. In the case that all habitats; mangrove, tidal marsh, and seagrass have availiable territory-level estimates, please refer to Table 2: IPCC Tier I Value Estimates for applicable global values.

```{r data setup, echo=FALSE}
tier1data <- allstocks %>% filter(TierIorII == "Tier I")
tier2data <- allstocks %>% filter(TierIorII == "Tier II")

```

```{r setuptier1, echo=FALSE, warning=FALSE, message=FALSE}
  if ("Tier I" %in% tiertable$Tier)
     {show_text = TRUE
     } else {show_text = FALSE}
```

```{r tier1eval, eval=show_text, echo=FALSE}
#tier 1 is always == IPCC global values 

#tibble of habitats in territory 
habitats <- allstocks %>% 
  select(territory, habitat) %>% 
  rename(Territory = territory,
         Habitat = habitat) %>% distinct()


#if no Tier I data listed, use IPCC estimates 
tier1stocks <- IPCC_tier_I %>% 
  mutate(Territory = chosen_geog) %>% 
  left_join(habitats) %>% 
  select(Territory, everything())


kable(tier1stocks, caption = "Tier I: Global Carbon Stock Estimates")
```

```{r tier1eval2, eval=show_text, echo=FALSE, results='asis'}
cat("Good news! The selected country has territory-level data for each available habitat. Please refer to Table 2: IPCC Tier I Value Estimates for applicable Tier I estimates and Table 5: Tier II Carbon Stock Estimates.")
```

### Tier II: Territory Level Carbon Stocks

This table includes country or territory level sediment carbon stock estimates for **`r chosen_geog`**. Estimates in this table were derived from available soil core data queried from the Coastal Carbon Atlas.

A selected country may have available territory-level values for one or multiple habitats, this is dependent on core data accessible through the Coastal Carbon Atlas. If a territory has habitat area, but that respective habitat type does not have an associated territory-level stock estimate, there is insufficient core data in the Coastal Carbon Atlas to estimate sediment carbon stock. At the time of analysis, referencing Version 1.5.0 of the Coastal Carbon Atlas and Data Library, **`r chosen_geog`** is represented by a total of `r cores` cores across `r habitats` distinct habitats.

```{r setuptier2_false, echo=FALSE, warning=FALSE, message=FALSE}
  if ("Tier II" %in% tiertable$Tier)
     {show_text = TRUE
     } else {show_text = FALSE}
```


```{r tier2, echo=FALSE}
library(knitr)
tier2stocks <- tier2data %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = soil_TierII_mean,
        `Lower CI` = soil_TierII_lowerCI,
        `Upper CI` = soil_TierII_upperCI,
        `Standard Error` = soil_TierII_se) %>%
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`, 3)) %>% 
  #janitor::remove_empty(which = "cols") %>% #removing any habitats without tier II data 
  dplyr::select(Country, Territory, Habitat, `Mean Stock (MgHa)`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(tier2stocks, caption = "Tier II: Territory-Level Carbon Stock Estimates") 
```

### Tier III Carbon Stocks

Tier III: modeled sediment carbon stocks were estimated, when available, using remote sensing data from Maxwell et al 2021 and Sanderman et al 2018. The table below details whether estimated values are available for any habitat in **`r chosen_geog`**, and details any overlap with associated global or territory-level estimates.

If there are no modeled estimates associated with the selected geography and specific habitat of interest, please refer to above tables of global and territory-level sediment carbon stocks. 

```{r tier 3setup1, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is tier3 data for at least one habitat 
tier3data <- allstocks %>% filter(!is.na(soil_TierIII_mean))

  if (chosen_geog %in% tier3data$territory)
     {show_text = TRUE
     } else {show_text = FALSE}

```

```{r tier overlap, echo=FALSE}
#tier 3 overlap with 1 and 2

#create variabletier3data#create variables to use in inline code
tierIII_gtlt_tier_II <- tier3data %>% select(tierIII_gtlt_tier_II)
tierII_overlaps_tierIII <- tier3data %>% select(tierII_overlaps_tierIII)  
tierIII_gtlt_tier_I <- tier3data %>% select(tierIII_gtlt_tier_I)
tierIII_overlaps_tierI <- tier3data %>% select(tierIII_overlaps_tierI)
```

```{r conditional_block1, eval=show_text, echo=FALSE, results='asis'}

tier3stocks <- tier3data %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = soil_TierIII_mean,
        `Lower CI` = soil_TierIII_lowerCI,
        `Upper CI` = soil_TierIII_upperCI) %>% 
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`,3)) %>% 
  select(Country, Territory, Habitat, `Mean Stock (MgHa)`, `Lower CI`, `Upper CI`)

 kable(tier3stocks, caption = "Tier III: Modeled Carbon Stock Estimates") 

#create variables to use in inline code
text1 <- tier3data %>% select(tierIII_gtlt_tier_II)
text2 <- tier3data %>% select(tierII_overlaps_tierIII)
text3  <- tier3data %>% select(tierIII_gtlt_tier_I)
text4 <- tier3data %>% select(tierIII_overlaps_tierI)
 
cat("In the selected geography, modeled estimated stock MgHa values are", text1, "territory-level values estimated from Coastal Carbon Atlas Data. In this case, the modeled", text2)

cat(". Additionally, estimated modeled stock values derived from remote sensing are", text3, "estimates global IPCC values. In this case, the modeled", text4, "estimate.")


#need to do this for each habitat 

#mangrove 
#text1mangrove <- text()


```

```{r tier3setup2, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is NO tier 3 data 
  if (!chosen_geog %in% tier3data$country)
     {show_text = TRUE
     } else {show_text = FALSE}

```

```{r conditional_block2, eval=show_text, echo=FALSE}
#this text prints if there is no available tier III data 
cat("There is currently no available modeled sediment carbon stock estimates for mangrove or marsh habitats in", chosen_geog)
```

```{r Tier3 Overlap, echo=FALSE}

#creating text and table for tier III overlap and gtlt tier I and II

# tier_overlap <- selectCountry(tier3data) %>% 
#   select(-c(stock_MgHa_mean, stock_MgHa_lowerCI, stock_MgHa_upperCI, tier))
# 
# kable(tier_overlap)
```
## Visualizations 

### 1. Data Metrics 
```{r data metrics, echo=FALSE}
map_input %>% filter(country == chosen_country) %>% 
  dplyr::count(carbon_pool, habitat, country) %>% 
  
  ggplot2::ggplot() + 
  geom_col(aes(habitat, n, fill = carbon_pool)) +
  # geom_errorbar(aes(x= habitat, ymin = cores, ymax = hectare_UpperCI, y= area_ha), width = 0.1) +
  coord_flip() +
  ylab("Number of Samples") + theme_bw(base_size = 20) +
  theme(legend.position = "bottom")

```

### 2. Emissions Factors 
```{r emissions factors, echo=FALSE, warning=FALSE}
allstocks %>% 
  select(-c(contains("gtlt"), contains("overlaps"), "TierIorII", "text_position")) %>%
  select(habitat, contains("stock"), contains("Tier")) %>% 
  select(-contains("Total")) %>% 
  pivot_longer(cols = -habitat, names_to = "tier", values_to = "stock") %>% 
  separate(tier, into = c("carbon_pool", "tier", "stat"), sep = "_") %>% 
  pivot_wider(id_cols = c("habitat", "carbon_pool", "tier"), names_from = stat, values_from = stock) %>% 
  
  ggplot2::ggplot(aes(mean, habitat, col = tier)) +
  # geom_boxplot(aes(stock_MgHa, habitat, col = `carbon pool`)) +
  geom_errorbar(aes(xmin = lowerCI, xmax  = upperCI), width = 0.1) +
  geom_point(size = 2, shape = 21, fill="white") +
  theme_bw() +
  facet_wrap(~`carbon_pool`, 
             # scales = "free", 
             dir = "v")+
  labs(title = "Emissions Factors: ADD TERRITORY")
  

```

### 3. Activity Data 
```{r, echo=FALSE, warning=FALSE}

allstocks %>% 
  ggplot2::ggplot() + 
  geom_col(aes(habitat, area_ha, fill = habitat)) +
  geom_errorbar(aes(x= habitat, ymin = area_ha_lowerCI, ymax = area_ha_upperCI, y= area_ha), width = 0.1) +
  coord_flip() +
  ylab("Area (hectares)") + theme_bw(base_size = 20) +
  labs(title = "Habitat Area")
  theme(legend.position = "bottom")
```


## Methods 
```{r, echo=FALSE, results='asis'}

cat(
"## Materials and Methodology 

This page details the data sources and methodology used to develop and estimate habitat areas and sediment carbon stocks, including data sources and term definitions. 

### Disclaimer

Country and territory names used in this map and dashboard have been aligned with current ESRI guidelines.The boundaries and names shown and the designations used on this map do not imply the expression of any opinion whatsoever on the part of the Coastal Carbon Network or the Smithsonian Institution concerning the legal status of any country, territory, city or area or of its authorities, or concerning the delimitation of its frontiers or boundaries. 

### Data Sources 

Multiple data sources were compiled in order to synthesize data for this tool. Worthington et al 2021 was used to estimate salt marsh area data, Bunting et al 2018 was used to estimate mangrove area data, and McKenzie et al 2021 was used to estimate seagrass area data, when available. Soil and vegetation data available in the Coastal Carbon Atlas, hosted by the Coastal Carbon Network, was used to estimate mean country-level sediment carbon stocks. This data was filtered to exclude cores shorter than 1 meter, in order to align with IPCC global carbon stock value standards. 

Modeled carbon stock values, estimated Tier III sediment carbon stocks, were estimated from Maxwell et al 2021 and Sanderman et al 2018 for all applicable territories. These estimated values were derived from remote sensing data present in these two syntheses, Maxwell et al 2021 for mangrove habitats, and Sanderman et al 2018 for tidal salt marsh habitats. 

When neither Tier II or Tier III values were available for a habitat in a geographic territory, the global IPCC sediment carbon stock value for that habitat is applied. 

### Mapping 

The interactive map featured in this tool was created using the leaflet package in R. OSM (default) and CartoDB provider tiles were used to create a global background layer. 

EEZ data source and Countries data source.
Spatial interpolation of unmonitored territories/watersheds. 

### Estimating Sediment Carbon Stocks

To estimate territory-level sediment carbon stocks, data from the Coastal Carbon Atlas and Data Library was used. 

### Definitions: Total Stocks 

Habitat Area (ha) : In hectares, area of specified habitat in selected territory.

Mean Stock (Mg/ha) : Mean soil carbon stock value in milligrams per hectare

Mean Stock Upper CI (Mg/ha) : Upper confidence interval for mean soil carbon stock value

Mean Stock Lower CI (Mg/ha): Lower confidence interval for mean soil carbon stock value 

CO2eq (TgC): (column name will be changed?) 

Reporting Tier: Level of reporting resolution for carbon stocks. Tiers listed include IPCC global values, country-level stock calculations, and modeled values. Tier I stocks correspond to IPCC global values, Tier II stocks correspond to country-level stock calculations, Tier III stocks correspond to modeled values. The reporting tier listed is the highest available resolution for a given territory and habitat. 

Reporting Insight: Specifies whether country-specific or modeled carbon stock averages overlap with IPCC global values")
```



## References
