---
title: "Summary Report: `r chosen_geog`"
output:
  html_document:
     toc: true
     toc_float: true
     theme: cerulean
bibliography: "`r here::here('app/data/bibs', bib)`"
engine: knitr
---

---
nocite: '@*'
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(kableExtra)
library(leaflet)
options(scipen = 999)

### This script writes the country insights pdf for the case of countries with NO AVAILABLE Tier II DATA 

#function to select only observations with chosen country 
selectGeography <- function(df){
  df %>% filter(territory == chosen_geog)
}

#country lookup table 
country_lookup <- read_csv("country_lookup.csv")

country <- country_lookup %>% 
  filter(territory == chosen_geog) %>% 
  select(country)
chosen_country <- as.character(country)

#function if needed to select only chosen country (CCA cores)
selectCountry <- function(df){
  df %>% filter(country == chosen_country)
}


#pull in curated app data, filtered to chosen territory 
app_data <- readRDS("app_data.rds") 
allstocks <- app_data$main_table %>% selectGeography()
totalstocks <- app_data$territory_tec %>% selectGeography()
map_input <- app_data$map_input #no territory listed 
map <- app_data$map_polys %>% selectGeography()

rm(app_data) #keeps the rds out of the environment 
```

```{r habitat areas, echo=FALSE}
#Area totals for individual habitats 
#mangrove values 
mangrove_area <- allstocks %>% filter(habitat == "mangrove")
area_ha_mangrove <- mangrove_area$area_ha
area_low_mangrove <- mangrove_area$area_ha_lowerCI
area_high_mangrove <- mangrove_area$area_ha_upperCI
#area_ha_mangrove <- 0

#marsh values
# marsh_area <- allstocks %>% filter(habitat == "marsh")
# area_ha_marsh <- marsh_area$area_ha
# area_low_marsh <- marsh_area$area_ha_lowerCI
# area_high_marsh <- marsh_area$area_ha_upperCI
area_ha_marsh <- 0

#seagrass values 
seagrass_area <- allstocks %>% filter(habitat == "seagrass")
area_ha_seagrass <- as.numeric(seagrass_area$area_ha)


```

```{r total stocks, echo=FALSE}
#total stocks
 countrystock <- totalstocks %>% selectGeography()

#total stocks// per total territory
stock_country <- countrystock$Total_Stocks
stock_country_low <- countrystock$Total_Stockers_LowerCI
stock_country_high <- countrystock$Total_Stockers_UpperCI
stock_country_se <- countrystock$Total_SE


#total stocks per habitat per territory 
mangrove_stocks <- allstocks %>% filter(habitat == "mangrove")
marsh_stocks <- allstocks %>% filter(habitat == "marsh")
seagrass_stocks <- allstocks %>% filter(habitat == "seagrass")


#mangrove
totalstocks_mangrove <- mangrove_stocks$Total_Stocks
totalstocks_mangrove_upper <- mangrove_stocks$area_ha_upperCI
totalstocks_mangrove_lower <- mangrove_stocks$area_ha_lowerCI

#marsh
# totalstocks_marsh <- marsh_stocks$Total_Stocks
# totalstocks_marsh_upper <- marsh_stocks$area_ha_upperCI
# totalstocks_marsh_lower <- marsh_stocks$Total_Stocks_LowerCI
totalstocks_marsh <- 0
   
#seagrass 
totalstocks_seagrass <- seagrass_stocks$Total_Stocks
totalstocks_seagrass_upper <- seagrass_stocks$area_ha_upperCI
totalstocks_seagrass_lower <- seagrass_stocks$Total_Stocks_LowerCI

```

## Welcome!

You have reached the CCN Inventory Tool Detailed Insights Report.

Great News! This geography has available data.

Potential data availability within the Inventory Tool ranges from Tier I, Tier II, or Tier III Carbon stock estimates. 

   - Tier I: Global sediment carbon stock estimates.

   - Tier II: Country- or territory-level sediment carbon stock estimates.

   - Tier III: Modeled country- or territory-level sediment carbon stock estimates. 

This is a downloadable document which includes specific insights and more detailed analysis for a selected territory or country, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass habitats. This report details information for the selected geography, **`r chosen_geog`**.

Explore the rest of the [inventory dashboard](https://ccrcn.shinyapps.io/BlueCarbonInventoryTool/) for more exciting visualizations, map features, and data.

If you have any questions or data you would like to add to the Coastal Carbon Network, please reach out to us at [CoastalCarbon\@si.edu](mailto:CoastalCarbon@si.edu){.email}.


### Map: `r chosen_geog`
```{r, echo=FALSE, warning=FALSE}
# "map" is created in set-up chunk using imported polygons 
# tier I <- no CCN samples

leaflet() %>%
  # basemap options
  addTiles(group = "OSM (default)") %>% 

  #add polygon layer for selected territory border
  addPolygons(data = map, weight = 2)


```
Figure 1. Interactive map of selected geography. If available for selected geography, vegetation and sediment samples included in the Coastal Carbon Atlas are displayed as marker points.

## Stocks Table 

Featured in the inventory tool dashboard, below is a table of any available sediment carbon stock estimates for the selected geography, **`r chosen_geog`**, for mangrove, tidal marsh, and seagrass habitat.

```{r main stocks table, echo=FALSE, error=FALSE, warning= FALSE}
main_table <- allstocks %>% 
  mutate(Continent = continent,
         Region = region,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (Mg/ha)` = soil_TierI_mean,
        `Reporting Tier` = TierIorII,
        `Area (hectares)` = signif(area_ha, 4)) %>% 
  select(Continent, Region, Territory, Habitat, `Area (hectares)`,`Mean Stock (Mg/ha)`, `Reporting Tier`)

kable(main_table)

```


## Total Carbon Stock Estimates

Total Carbon stock estimates were calculated for each geography in total, and for mangrove, tidal marsh, and seagrass habitats, if applicable. At this time total Carbon stock estimates for each country do not include seagrass estimates due to the lack of a global, transparent, and independently assessed seagrass habitat map. However, best available areas and stocks for **`r chosen_geog`** are explored in the following 'Wetland Areas and Activities' section.

### Country Level Total Stock Estimates

We estimate that **`r chosen_geog`** contains approximately `r signif(stock_country,4)` megagrams per hectare (Mg/ha) soil carbon to a depth of 1 meter. 


```{r table1, echo=FALSE, warning=FALSE}

totals <- allstocks %>% 
  mutate(Country = country,
         Territory = territory,
         Habitat = habitat,
         `Total Stock` = signif(Total_Stocks, 4),
         `Upper CI` = signif(Total_Stocks_UpperCI, 4),
         `Lower CI` = signif(Total_Stocks_LowerCI, 4),
         `Standard Error` = signif(Total_Stocks_se, 4)) %>% 
  select(Territory, Country, Habitat, `Total Stock`, `Upper CI`, `Lower CI`,
         `Standard Error`)

# #if NA then 0?// no data available in cca syntheses 
# totalstocks_mangrove[is.na(totalstocks_mangrove)] <- 0
# totalstocks_marsh[is.na(totalstocks_marsh)] <- 0
# totalstocks_seagrass[is.na(totalstocks_seagrass)] <- 0

kable(totals, caption = "Total Territory-Level Stocks")
```

### Mangrove Estimates

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of **`r signif(totalstocks_mangrove)`** megagrams per hectare (Mg/ha) soil carbon in mangrove habitats.

### Tidal Marsh Estimates

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of **`r signif(totalstocks_marsh)`** megagrams per hectare (Mg/ha) soil carbon in tidal marsh habitats.

### Seagrass Estimates

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of **`r signif(totalstocks_seagrass, 3)`** megagrams per hectare (Mg/ha) soil carbon in seagrass habitats.

## Wetland Areas and Activities

### Mangrove Area

```{r mangrovearea, echo=FALSE, warning=FALSE, message=FALSE}
#errors with zeros, evaluate if area is greater than numeric zero 
# area_ha_mangrove_eval <- isTRUE(area_ha_mangrove > 0)
# 
# #evaluating the case of available mangrove area 
#   if (isTRUE(area_ha_mangrove_eval))
#      {show_text = TRUE
#   } else {show_text = FALSE}
# 
#   if (area_ha_mangrove_eval == FALSE)
#      {show_text2 = TRUE
#   } else {show_text2 = FALSE}

```

```{r mangroveeval, echo=FALSE, results='asis'}

# cat("We estimate mangrove area in",chosen_geog, "to be", signif(area_low_mangrove, 3), "to", signif(area_high_mangrove, 3),"hectares with a mean estimate of", signif(area_ha_mangrove, 3),"hectares according to Global Mangrove Watch, Bunting et al 2018.")
```

```{r mangroveeval2, echo=FALSE, results='asis'}
# cat("We estimate mangrove area in", chosen_geog, "to be approximately 0 hectares according to Global Mangrove Watch, Bunting_et_al_2018.")
```
We estimate mangrove area in `r chosen_geog` to be approximately 0 hectares according to Global Mangrove Watch, Bunting et al 2018.

### Tidal Marsh Area

```{r marsharea, echo=FALSE, warning=FALSE, message=FALSE}

# area_ha_marsh_eval <- isTRUE(area_ha_marsh > 0)
# 
# #evaluating the case of available mangrove area 
#   if (isTRUE(area_ha_marsh_eval))
#      {show_text = TRUE
#   } else {show_text = FALSE}
# 
#   if (area_ha_marsh_eval == FALSE)
#      {show_text2 = TRUE
#   } else {show_text2 = FALSE}


```

```{r marsheval, echo=FALSE, results='asis'}

# area_marsh <- as.numeric(area_ha_marsh) %>% signif(3)
# area_low_marsh <- as.numeric(area_low_marsh) %>% signif(3)
# area_high_marsh <- as.numeric(area_high_marsh) %>% signif(3)
# 
# cat("We estimate tidal marsh area in",chosen_geog, "to be", area_low_marsh, "to", area_high_marsh,"hectares with a mean estimate of", signif(area_ha_marsh, 3),"hectares according to Worthington et al 2024.") 
```

```{r marsheval2, echo=FALSE, results='asis'}
# cat("We estimate tidal marsh area in", chosen_geog, "to be approximately 0 hectares according to Worthington et al 2024.")
```

We estimate tidal marsh area in `r chosen_geog` to be approximately 0 hectares according to Global Mangrove Watch, Bunting et al 2018.


### Seagrass Area

```{r seagrassarea, echo=FALSE, warning=FALSE, message=FALSE}

area_ha_seagrass_eval <- isTRUE(area_ha_seagrass > 0)

#evaluating the case of available mangrove area 
  if (isTRUE(area_ha_seagrass_eval))
     {show_text = TRUE
  } else {show_text = FALSE}

  if (area_ha_seagrass_eval == FALSE)
     {show_text2 = TRUE
  } else {show_text2 = FALSE}


```

```{r seagrasseval, echo=FALSE, results='asis'}

#area_seagrass <- as.numeric(area_ha_seagrass) %>% signif(3)

#seagrass_high_confidence <- 1 - 100
#seagrass_low_confidence <- 1 - 100

# cat("We estimate seagrass area in",chosen_geog, "to have a mean estimate of", signif(area_ha_seagrass, 3),"hectares according to McKenzie et al 2020, which aggregates data from multiple sources.")


# cat("McKenzie et al 2020 classifies seagrass area estimates as either high or medium to low confidence. seagrass_high_confidence % of the estimated seagrass area of",chosen_geog, "is considered high to medium confidence, while seagrass_low_confidence % of the estimated seagrass area is categorized as low confidence.")
```

```{r seagrasseval2, echo=FALSE, results='asis'}
#cat("We estimate seagrass area in", chosen_geog, "to be approximately 0 hectares according to McKenzie et al 2020.")
```
We estimate seagrass area in `r chosen_geog` to have a mean estimate of `r signif(area_ha_seagrass, 3) ` hectares according to McKenzie et al 2020.

## Calculated Stocks and Emissions Factors

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_geog`**.

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas.

Country-specific soil cores from **`r chosen_geog`** have not yet been included in the Coastal Carbon Atlas. In this case, we have instead applied the IPCC Tier I estimated value. IPCC Tier I estimates for mangrove, marsh, and seagrass habitats are listed below.

### Tier I Carbon Stocks: Global Values

```{r IPCC, echo=FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- tibble(Habitat = c("mangrove", "marsh", "seagrass"),
                      Mean = c(386, 255, 108),
                      `Lower CI` = c(351,254,84),
                      `Upper CI` = c(424,297,139))

kable(IPCC_tier_I, caption = "IPCC Tier I: Global Carbon Stock Estimates")
```


```{r tier1, echo=FALSE}

# tier1data <- allstocks %>% 
#   select(territory, country, habitat, soil_TierI_mean, soil_TierI_lowerCI, soil_TierI_upperCI)
# 
# tier1stocks <- tier1data %>% 
#   rename(Country = country,
#          Territory = territory,
#          Habitat = habitat,
#         `Mean Stock (MgHa)` = soil_TierI_mean,
#         `Lower CI` = soil_TierI_lowerCI,
#         `Upper CI` = soil_TierI_upperCI)
# 
# 
# kable(tier1stocks)
```

### Tier II Carbon Stocks: Territory-Level Values

Tier II, or territory-level carbon stocks were estimated, when available, from soil core data currently housed in the Coastal Carbon Atlas.

Territory-Level Data Availability for **`r chosen_geog`**: The selected geography, **`r chosen_geog`** does not currently have available Tier II carbon stock estimates at a depth of 1 meter. Please refer to IPCC global values. 

### Tier III Carbon Stocks: Modeled Values

Tier III, or modeled carbon stocks were estimated, when available, from remote sensing data from Maxwell et al 2021 for mangrove habitats, and Sanderman et al 2018 for tidal marsh habitats. The following details whether estimated values are available for **`r chosen_geog`**, and any overlap with associated Tier I or Tier II values.

If there are no modeled carbon stock estimates associated with the selected country, please refer to Tier I and/or Tier II tables.

Modeled Data Availability for **`r chosen_geog`**: There are currently no Tier III values for **`r chosen_geog`** in the CCN Inventory Dashboard. Please refer to IPCC global values.

## Visualizations 

<!-- ### 1. Data Metrics  -->
<!-- ```{r data metrics, echo=FALSE} -->
<!-- map_input %>% filter(country == chosen_country) %>%  -->
<!--   dplyr::count(carbon_pool, habitat, country) %>%  -->

<!--   ggplot2::ggplot() +  -->
<!--   geom_col(aes(habitat, n, fill = carbon_pool)) + -->
<!--   # geom_errorbar(aes(x= habitat, ymin = cores, ymax = hectare_UpperCI, y= area_ha), width = 0.1) + -->
<!--   coord_flip() + -->
<!--   ylab("Number of Samples") + theme_bw(base_size = 20) + -->
<!--   theme(legend.position = "bottom") -->

<!-- ``` -->

### 2. Emissions Factors 
```{r emissions factors, echo=FALSE, warning=FALSE}
allstocks %>% 
  plot_ly(x = ~habitat, y = ~soil_TierI_mean, type = "bar", 
           error_y = ~list(array = soil_TierI_upperCI - soil_TierI_mean, 
                           arrayminus = soil_TierI_mean - soil_TierI_lowerCI, color = "black"),
           name = "IPCC global value") %>% 
     # add_trace(y = ~soil_TierII_mean, 
     #           error_y = ~list(array = soil_TierII_upperCI - soil_TierII_mean, 
     #                           arrayminus = soil_TierII_mean - soil_TierII_lowerCI, color = "black"),
     #           name = "Country-specific value") %>% 
     # add_trace(y = ~soil_TierIII_mean, 
     #           error_y = ~list(array = soil_TierIII_upperCI - soil_TierIII_mean, 
     #                           arrayminus = soil_TierIII_mean - soil_TierIII_lowerCI, color = "black"),
     #           name = "Modeled value") %>% 
     layout(
       title = "Tier 1: IPCC Global Values",
       xaxis = list(title = "Habitat Type"),
       yaxis = list(title = "Soil Carbon Stock (Mg/ha)"))

```

### 3. Activity Data 
```{r, echo=FALSE, warning=FALSE}
allstocks %>% 
  plot_ly(x = ~habitat, y = ~area_ha, type = "bar",
              # color = ~habitat, colors = "Set1",
              error_y = ~list(array = area_ha_upperCI - area_ha, 
                              arrayminus = area_ha - area_ha_lowerCI, color = "black")) %>% 
      layout(
        title = "Habitat Area (ha)",
        xaxis = list(title = "Habitat Type"),
        yaxis = list(title = "Area (ha)"))
```



## References
