---
title: "Summary Report: `r chosen_geog`"
output:
  html_document:
     toc: true
     toc_float: true
     theme: cerulean
bibliography: "`r here::here('app/data/bibs', bib)`"
engine: knitr
---
---
nocite: '@*'
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(kableExtra)
library(leaflet)
library(plotly)
library(sf)
options(scipen = 999)

### This script writes the country insights pdf for the case of countries with AVAILABLE Tier II DATA 

#function to select only observations with chosen country 
selectGeography <- function(df){
  df %>% dplyr::filter(territory == chosen_geog)
}

#country lookup table 
country_lookup <- read_csv("country_lookup.csv")

country <- country_lookup %>%
  filter(territory == chosen_geog) %>%
  select(country)
chosen_country <- as.character(country)

#function if needed to select only chosen country (CCA cores)
selectCountry <- function(df){
  df %>% filter(country == chosen_country)
}

#pull in curated app data, filtered to chosen territory 
app_data <- readRDS("app_data.rds") 
allstocks <- app_data$main_table %>% selectGeography()
totalstocks <- app_data$territory_tec %>% selectGeography()
map_input <- app_data$map_input %>% selectGeography
#map <- app_data$map_polys %>% selectGeography()
map <- st_read("territory_shapefiles/ccn_territory_map.shp") %>% selectGeography()

rm(app_data) #keeps the rds out of the environment 
```

```{r habitat areas, echo=FALSE}
#Area totals for individual habitats 

#mangrove values 
mangrove_area <- allstocks %>% filter(habitat == "mangrove")
area_ha_mangrove <- signif(mangrove_area$area_ha, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
area_low_mangrove <- signif(mangrove_area$area_ha_lowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
area_high_mangrove <- signif(mangrove_area$area_ha_upperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)

#marsh values
marsh_area <- allstocks %>% filter(habitat == "marsh")
area_ha_marsh <- signif(marsh_area$area_ha, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
area_low_marsh <- signif(marsh_area$area_ha_lowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
area_high_marsh <- signif(marsh_area$area_ha_upperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)


#seagrass values 
seagrass_area <- allstocks %>% filter(habitat == "seagrass")
area_ha_seagrass <- signif(seagrass_area$area_ha, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
#no CIs
```

```{r total stocks, echo=FALSE}
#total stocks
 countrystock <- totalstocks %>% selectGeography()
  #selectCountry()

#total stocks// per total territory
stock_country <- signif(countrystock$Total_Stocks, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
stock_country_low <- signif(countrystock$Total_Stockers_LowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
stock_country_high <- signif(countrystock$Total_Stockers_UpperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
stock_country_se <- signif(countrystock$Total_SE, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)

#total stocks per habitat per territory 
#this is now taken from different table 
mangrove_stocks <- allstocks %>% filter(habitat == "mangrove")
marsh_stocks <- allstocks %>% filter(habitat == "marsh")
seagrass_stocks <- allstocks %>% filter(habitat == "seagrass")

#mangrove
totalstocks_mangrove <- signif(mangrove_stocks$Total_Stocks, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_mangrove_upper <- signif(mangrove_stocks$area_ha_upperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_mangrove_lower <- signif(mangrove_stocks$area_ha_lowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
  
#marsh
totalstocks_marsh <- signif(marsh_stocks$Total_Stocks, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_marsh_upper <- signif(marsh_stocks$area_ha_upperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_marsh_lower <- signif(marsh_stocks$Total_Stocks_LowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)


#seagrass 
totalstocks_seagrass <- signif(seagrass_stocks$Total_Stocks, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_seagrass_upper <- signif(seagrass_stocks$area_ha_upperCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)
totalstocks_seagrass_lower <- signif(seagrass_stocks$Total_Stocks_LowerCI, 4) %>% prettyNum(big.mark = ",", scientific = FALSE)

```

## Welcome!

You have reached the CCN Inventory Tool Detailed Insights Report.

Congratulations! This geography has available data.

Potential data availability within the Inventory Tool ranges from Tier I, Tier II, or Tier III Carbon stock estimates. 

   - Tier I: Global sediment carbon stock estimates.

   - Tier II: Country- or territory-level sediment carbon stock estimates.

   - Tier III: Modeled country- or territory-level sediment carbon stock estimates. 


This document includes country-specific insights and more detailed analysis, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass habitats. This report details information for the selected geography, **`r chosen_geog`**.

Explore the rest of the dashboard for more exciting visualizations, map features and data.

If you have any questions or data you would like to add to the Coastal Carbon Network, please reach out to us at [CoastalCarbon\@si.edu](mailto:CoastalCarbon@si.edu){.email}.

```{r, echo=FALSE, warning=FALSE}
# "map" is created in set-up chunk using imported polygons 
# tier II, CCA samples overlay 

chosen_map_input <- map_input %>% selectGeography()

leaflet() %>%
  # basemap options
  addTiles(group = "OSM (default)") %>% 

  #add polygon layer for selected territory border
  addPolygons(data = map, weight = 2) %>% 
   # add data points (global cca samples)
        addCircleMarkers(data = chosen_map_input,
                         lng = ~longitude, lat = ~latitude, radius = 2,
                         label = ~paste(habitat, carbon_pool, "data", sep = " "),
                         group = "Samples", 
                         # eventually have veg be plotted separately for color coding
                         color = "green") 


```
This map represents the area of **`r chosen_geog`** and the locations of both soil cores and any vegetation samples included in the Coastal Carbon Atlas used for the estimation of carbon stock.


## Total Carbon Stock Estimates

Total Carbon stock estimates were calculated for each geography in total, and for mangrove, tidal marsh, and seagrass habitats, when available.

We estimate that **`r chosen_geog`** contains a mean estimate of `r stock_country` megagrams sediment carbon stock. 

We estimate that **`r chosen_geog`** contains between `r stock_country_low` to `r stock_country_high` megagrams sediment carbon to a depth of 1 m, with a mean estimate of `r stock_country` megagrams.

### Total Territory-Level Stocks
```{r table1 territory total, echo=FALSE}

totals <- totalstocks %>%
  selectGeography() %>%
  mutate(Country = country,
         Territory = territory,
         `Total Stocks` = signif(Total_Stocks, 4),
         `Upper CI` = signif(Total_Stockers_UpperCI, 4),
         `Lower CI` = signif(Total_Stockers_LowerCI, 4),
         `Total Stocks` = prettyNum(big.mark = ",", scientific = FALSE)) %>%
  select(Country, Territory, `Total Stocks`, `Upper CI`, `Lower CI`)

kable(totals, caption = "Total Geography Level Carbon Stocks",
      format.args = list(big.mark = ","))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
# #decide if mangrove data 
# #errors with zeros, evaluate if area is greater than numeric zero 
# mangrove_data_eval <- isTRUE(totalstocks_mangrove > 0)
# 
# #evaluating the case of available mangrove area 
#   if (isTRUE(mangrove_data_eval))
#      {show_textmangrove = TRUE
#   } else {show_textmangrove = FALSE}
# 
#   if (mangrove_data_eval == FALSE)
#      {show_text2mangrove = TRUE
#   } else {show_text2mangrove = FALSE}

```

### Total Mangrove Stocks 
```{r table2 mangrove, echo=FALSE}
totalstocks_mangrove <- totalstocks_mangrove


# cat("This total estimate includes total mangrove soil carbon stocks, with a mean estimate of", totalstocks_mangrove, "megagrams per hectare soil carbon to a depth of 1 meter.")


mangrove_totals <- allstocks %>% filter(habitat == "mangrove") %>% 
     mutate(Country = country,
            Territory = territory,
            Habitat = habitat,
            Habitat = "Mangrove",
            `Total Stock` = signif(Total_Stocks, 4),
            `Lower CI` = signif(Total_Stocks_LowerCI, 4),
            `Upper CI` = signif(Total_Stocks_UpperCI, 4),
            `Standard Error` = signif(Total_Stocks_se, 4)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(Country, Territory, Habitat, `Total Stock`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(mangrove_totals, 
      format.args = list(big.mark = ","))
```

This total estimate includes total mangrove soil carbon stocks, with a mean estimate of `r totalstocks_mangrove` megagrams per hectare soil carbon to a depth of 1 meter.

```{r, echo=FALSE, warning=FALSE, message=FALSE} 
#decide if marsh data 

#decide if mangrove data, only print data if stocks are present 
#errors with zeros, evaluate if area is greater than numeric zero 
marsh_data_eval <- isTRUE(totalstocks_marsh > 0)

#evaluating the case of available mangrove area 
  if (isTRUE(marsh_data_eval))
     {show_textmarsh = TRUE
  } else {show_textmarsh = FALSE}

  if (marsh_data_eval == FALSE)
     {show_text2marsh = TRUE
  } else {show_text2marsh = FALSE}


```

### Total Marsh Stocks 
```{r table3 marsh, eval=show_textmarsh, echo=FALSE, results='asis'}

# totalstocks_marsh_lower <- signif(totalstocks_marsh_lower, 4)
# totalstocks_marsh_upper <- signif(totalstocks_marsh_upper, 4)
# totalstocks_marsh <- signif(totalstocks_marsh, 4)

# cat("This total estimate also includes total tidal marsh carbon stocks, ranging from", totalstocks_marsh_lower, "to", totalstocks_marsh_upper,", with a mean estimate of",totalstocks_marsh,"megagrams per hectare.")


marsh_totals <- allstocks %>% filter(habitat == "marsh") %>% 
     mutate(Country = country,
            Territory = territory,
            Habitat = habitat,
            Habitat = "Marsh",
            `Total Stock` = signif(Total_Stocks, 4),
            `Lower CI` = signif(Total_Stocks_LowerCI, 4),
            `Upper CI` = signif(Total_Stocks_UpperCI, 4),
            `Standard Error` = signif(Total_Stocks_se, 4)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(Country, Territory, Habitat, `Total Stock`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(marsh_totals,
      format.args = list(big.mark = ","))
```

This total estimate also includes total tidal marsh carbon stocks, with a mean estimate of `r totalstocks_marsh` megagrams soil carbon to a depth of 1 meter.


```{r, echo=FALSE, warning=FALSE, message=FALSE} 
# #decide if seagrass data 
# 
# #only print table if stocks are present 
# #errors with zeros, evaluate if area is greater than numeric zero 
# seagrass_data_eval <- isTRUE(totalstocks_seagrass > 0)
# 
# #evaluating the case of available mangrove area 
#   if (isTRUE(seagrass_data_eval))
#      {show_textseagrass = TRUE
#   } else {show_textseagrass = FALSE}
# 
#   if (seagrass_data_eval == FALSE)
#      {show_text2seagrass = TRUE
#   } else {show_text2seagrass = FALSE}

```

### Total Seagrass Stocks
```{r table4 seagrass, echo=FALSE}

#totalstocks_seagrass <- signif(totalstocks_seagrass, 4)

# cat("We estimate seagrass soil carbon stocks for", chosen_geog, "to have a mean estimate of", totalstocks_seagrass, "megagrams per hectare sediment carbon stock.")


seagrass_totals <- allstocks %>% filter(habitat == "seagrass") %>%
     mutate(Territory = territory,
            Habitat = habitat,
            Habitat = "Seagrass",
            `Total Stock` = signif(Total_Stocks, 4)) %>% 
  #no CI columns or se for seagrass total stocks 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data
  select(Territory, Habitat, `Total Stock`)


kable(seagrass_totals,
      format.args = list(big.mark = ","))
```

We estimate seagrass habitats in `r chosen_geog` to have a mean of `r totalstocks_seagrass` megagrams of sediment carbon stock. This value is not included in the total country-level stock estimate due to the lack of a global, transparent, and independently assessed seagrass habitat map.

## Wetland Areas and Activities

We estimate mangrove area in **`r chosen_geog`** to be `r area_low_mangrove` to `r area_high_mangrove` hectares, with a mean estimate of `r area_ha_mangrove` hectares according to Global Mangrove Watch, @Bunting_et_al_2018.

We estimate tidal marsh area in **`r chosen_geog`** to be `r area_low_marsh` to `r area_high_marsh` hectares, with a mean estimate of     `r area_ha_marsh` according to @Worthington_et_al_2024.

We estimate seagrass area to be **`r chosen_geog`** to be a mean of `r area_ha_seagrass` hectares, according to @McKenzie_et_al_2020, which aggregates global seagrass data from a number of sources.

<!-- @McKenzie_et_al_2020 classifies seagrass area estimates as either high or medium-to-low confidence. seagrass_area_high_confidence % of the estimated seagrass area of **`r chosen_geog`** is considered high to medium confidence, while seagrass_area_low_confidence % of the estimated seagrass area is categorized as low confidence. -->

## Calculated Stocks and Emissions Factors

```{r N_cores, echo = FALSE, warning=FALSE}
#cores, sites, habitats for soil cores included in tier II calcs 
#cores <- chosen_map_input %>% filter(carbon_pool == "soil") %>% count() 
#veg_samples <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% count()
sites <- chosen_map_input %>% filter(carbon_pool == "soil") %>% select(site_id) %>% distinct() %>% count()
#veg_sites <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% select(site_id) %>% distinct() %>% count()
habitats <- chosen_map_input %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct() %>% count()
#veg_habitat <- chosen_map_input %>% filter(carbon_pool == "vegetation") %>% select(habitat) %>% distinct() %>% count()

habitat.list <- selectCountry(map_input) %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct()
#veg.habitat.list <- selectCountry(map_input) %>% filter(carbon_pool == "vegetation") %>% select(habitat) %>% distinct()
habitat.list <- as.character(habitat.list)
#veg.habitat.list <- as.character(veg.habitat.list)


## veg data present? 
#veg_data_present <- if_else("vegetation" %in% chosen_map_input$carbon_pool, "also contains", "does not contain")


```

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_geog`**.

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas.

Within the Coastal Carbon Atlas, data from the selected geography, **`r chosen_geog`**, includes soil samples but does not include vegetation samples

Soil core data from **`r chosen_geog`** included in the Coastal Carbon Atlas consists of 1107 soil profiles from 540 unique sites. This data comes from 5 distinct habitat types, including marsh, mangrove, seagrass, scrub/shrub and swamp.


```{r if veg data, echo=FALSE}

# veg_data_present
# #if vegetation data 
#  if (veg_data_present == "also contains")
#      {show_text = TRUE
#   } else {show_text = FALSE}

```

```{r veg data,  echo=FALSE, results='asis'}

# cat("Vegetation data included in the Coastal Carbon Atlas consists of", veg_samples, "samples from", veg_sites, "unique sites.")
# 
# cat("This data has been sourced from", veg_habitat, "which includes", veg.habitat.list)

```

In the case that there is not yet any country specific information in the Coastal Carbon Atlas, we instead applied IPCC Tier I estimate, which are estimated global values. IPCC Global estimates for mangrove, marsh, and seagrass ecosystems are listed below.

The table titled "Availiability of Territory-Level Data" in this section also details whether the calculated Tier II, or territory-level sediment carbon stock value is significantly different from the estimated global values. This is observed in the "Reporting Insight" column, and can be helpful in observing when a habitat has a higher or lower estimated sediment carbon stock value than the IPCC Global estimates. 

```{r IPCC, echo =FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- as_tibble(data.frame(Habitat = c("mangrove", "marsh", "seagrass"),
                                      Mean = c(386, 255, 108),
                                      `Lower_CI` = c(351,254,84),
                                      `Upper_CI` = c(424,297,139)))

kable(IPCC_tier_I, caption = "IPCC Global Value Estimates (Mg per hectare)")
```

## Data Availiability by Tier - Reporting Insight

```{r tier availability, echo = FALSE}
#create table of available tier estimates and any overlap 

tier_mangrove <- allstocks %>% filter(habitat == "mangrove") %>% select(TierIorII)
tier_marsh <- allstocks %>% filter(habitat == "marsh") %>% select(TierIorII)
tier_seagrass <- allstocks %>% filter(habitat == "seagrass") %>% select(TierIorII)

tiertable <- allstocks %>% select(country, territory, habitat, TierIorII, tier_II_overlaps_TierI) %>% 
  dplyr::rename(Country = country, 
         Territory = territory,
         Habitat = habitat,
         Tier = TierIorII,
         `Reporting Insight` = tier_II_overlaps_TierI) %>% 
  mutate(`Reporting Insight` = case_when(is.na(`Reporting Insight`) ~ "No Tier II data available at this time",
                                         grepl("overlap", `Reporting Insight`) ~ "Territory-level average overlaps global value",
                                         grepl("significantly less", `Reporting Insight`) ~ 
                                           "Territory-level average is significantly less than global value",
                                         grepl("significantly greater", `Reporting Insight`) ~ 
                                           "Territory-level average is significantly greater than global value",
                                         TRUE ~ `Reporting Insight`),
        Tier = if_else(Tier == "Tier II", "Territory-Level", "Global"))

kable(tiertable, caption = "Availiability of Territory-Level Data")

```

### Tier I Carbon Stocks - Global Estimates 

This section includes estimated global sediment carbon stocks included for **`r chosen_geog`**. In the case that any habitats; mangrove, tidal marsh, or seagrass do not have availiable territory-level estimates, please refer to Table 2: IPCC Tier I Value Estimates for applicable global values.

Good news! The selected country has territory-level data for each available habitat. Please refer to Table 2: IPCC Tier I Value Estimates for applicable Tier I estimates and Table 5: Tier II Carbon Stock Estimates."

```{r data setup, echo=FALSE}
tier1data <- allstocks %>% filter(TierIorII == "Tier I")
tier2data <- allstocks %>% filter(TierIorII == "Tier II")

```

```{r setuptier1, echo=FALSE, warning=FALSE, message=FALSE}
  # if ("Tier I" %in% tiertable$Tier)
  #    {show_text = TRUE
  #    } else {show_text = FALSE}
```

```{r tier1eval, , echo=FALSE}
# #tier 1 is always == IPCC global values 
# 
# #tibble of habitats in territory 
# habitats <- allstocks %>% 
#   select(territory, habitat) %>% 
#   rename(Territory = territory,
#          Habitat = habitat) %>% distinct()
# 
# 
# #if no Tier I data listed, use IPCC estimates 
# tier1stocks <- IPCC_tier_I %>% 
#   mutate(Territory = chosen_geog) %>% 
#   left_join(habitats) %>% 
#   select(Territory, everything())
# 
# 
# kable(tier1stocks, caption = "Tier I: Global Carbon Stock Estimates")
```

```{r tier1eval2,  echo=FALSE}
# cat("Good news! The selected country has territory-level data for each available habitat. Please refer to Table 2: IPCC Tier I Value Estimates for applicable Tier I estimates and Table 5: Tier II Carbon Stock Estimates.")
```

### Tier II: Territory Level Carbon Stocks

This table includes country or territory level sediment carbon stock estimates for **`r chosen_geog`**. Estimates in this table were derived from available soil core data queried from the Coastal Carbon Atlas.

A selected country may have available territory-level values for one or multiple habitats, this is dependent on core data accessible through the Coastal Carbon Atlas. If a territory has habitat area, but that respective habitat type does not have an associated territory-level stock estimate, there is insufficient core data in the Coastal Carbon Atlas to estimate sediment carbon stock. At the time of analysis, referencing Version 1.5.0 of the Coastal Carbon Atlas and Data Library, **`r chosen_geog`** is represented by a total of 1107 cores across 5 distinct habitat(s).

```{r setuptier2_false, echo=FALSE, warning=FALSE, message=FALSE}
  # if ("Tier II" %in% tiertable$Tier)
  #    {show_text = TRUE
  #    } else {show_text = FALSE}
```


```{r tier2, echo=FALSE}
library(knitr)
tier2stocks <- tier2data %>% 
  dplyr::rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = soil_TierII_mean,
        `Lower CI` = soil_TierII_lowerCI,
        `Upper CI` = soil_TierII_upperCI,
        `Standard Error` = soil_TierII_se) %>%
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`, 3),
         `Standard Error` = signif(`Standard Error`, 3)) %>% 
  #janitor::remove_empty(which = "cols") %>% #removing any habitats without tier II data 
  dplyr::select(Country, Territory, Habitat, `Mean Stock (MgHa)`, `Lower CI`, `Upper CI`, `Standard Error`)

kable(tier2stocks, caption = "Tier II: Territory-Level Carbon Stock Estimates",
      format.args = list(big.mark = ",")) 
```

### Tier III Carbon Stocks

Tier III: modeled sediment carbon stocks were estimated, when available, using remote sensing data from Maxwell et al 2021 and Sanderman et al 2018. The table below details whether estimated values are available for any habitat in **`r chosen_geog`**, and details any overlap with associated global or territory-level estimates.

If there are no modeled estimates associated with the selected geography and specific habitat of interest, please refer to above tables of global and territory-level sediment carbon stocks. 

```{r tier 3setup1, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is tier3 data for at least one habitat 
tier3data <- allstocks %>% filter(!is.na(soil_TierIII_mean))

  # if (chosen_geog %in% tier3data$territory)
  #    {show_text = TRUE
  #    } else {show_text = FALSE}

```

```{r tier overlap, echo=FALSE}
#tier 3 overlap with 1 and 2

marshtier3 <- tier3data %>% filter(habitat == "marsh")
mangrovetier3 <- tier3data %>% filter(habitat == "mangrove")

#create variabletier3data#create variables to use in inline code
tierIII_gtlt_tier_II_mangrove <- mangrovetier3 %>%  select(tierIII_gtlt_tier_II) 
tierII_overlaps_tierIII_mangrove <- mangrovetier3 %>% select(tierII_overlaps_tierIII)  
tierIII_gtlt_tier_I_mangrove <- mangrovetier3 %>% select(tierIII_gtlt_tier_I)
tierIII_overlaps_tierI_mangrove <- mangrovetier3 %>% select(tierIII_overlaps_tierI)

tierIII_gtlt_tier_II_marsh <- marshtier3 %>% select(tierIII_gtlt_tier_II)
tierII_overlaps_tierIII_marsh <- marshtier3 %>% select(tierII_overlaps_tierIII)  
tierIII_gtlt_tier_I_marsh <- marshtier3 %>% select(tierIII_gtlt_tier_I)
tierIII_overlaps_tierI_marsh <- marshtier3 %>% select(tierIII_overlaps_tierI)

```

```{r conditional_block1, echo=FALSE}

tier3stocks <- tier3data %>% 
  dplyr::rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = soil_TierIII_mean,
        `Lower CI` = soil_TierIII_lowerCI,
        `Upper CI` = soil_TierIII_upperCI) %>% 
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`,3)) %>% 
  select(Country, Territory, Habitat, `Mean Stock (MgHa)`, `Lower CI`, `Upper CI`)

 kable(tier3stocks, caption = "Tier III: Modeled Carbon Stock Estimates",
       format.args = list(big.mark = ",")) 

#create variables to use in inline code
text1marsh <- tierIII_gtlt_tier_II_marsh
text2marsh <- tierII_overlaps_tierIII_marsh
text3marsh  <- tierIII_gtlt_tier_I_marsh
text4marsh <- tierIII_overlaps_tierI_marsh

text1mangrove <- tierIII_gtlt_tier_II_mangrove
text2mangrove <- tierII_overlaps_tierIII_mangrove
text3mangrove  <- tierIII_gtlt_tier_I_mangrove
text4mangrove <- tierIII_overlaps_tierI_mangrove

# In the selected geography, modeled estimated stock MgHa values for marsh habitats are, `r text1marsh` territory-level values estimated from Coastal Carbon Atlas Data. In this case, the modeled `r text2marsh`. Additionally, estimated modeled stock values derived from remote sensing are `r text3marsh` estimates global IPCC values. In this case, the modeled `r text4marsh` global estimates.
# 
# In mangrove habitats, modeled estimated stock MgHa values are, `r text1mangrove` territory-level values estimated from Coastal Carbon Atlas Data. In this case, the modeled `r text2mangrove`. Additionally, estimated modeled stock values derived from remote sensing are `r text3mangrove` estimates global IPCC values. In this case, the modeled `r text4mangrove` global estimates.
# 

```

In the selected geography, `r chosen_geog`, modeled estimated stock MgHa values for marsh habitats are, `r text1marsh` territory-level values estimated from Coastal Carbon Atlas Data. In this case, the modeled `r text2marsh`. Additionally, estimated modeled stock values derived from remote sensing are `r text3marsh` estimates global IPCC values. In this case, the modeled `r text4marsh` global estimates.

While in mangrove habitats, modeled estimated stock MgHa values are, `r text1mangrove` territory-level values estimated from Coastal Carbon Atlas Data. In this case, the modeled `r text2mangrove`. Additionally, estimated modeled stock values derived from remote sensing are `r text3mangrove` estimates global IPCC values. In this case, the modeled `r text4mangrove` global estimates.


```{r tier3setup2, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is NO tier 3 data 
  # if (!chosen_geog %in% tier3data$country)
  #    {show_text = TRUE
  #    } else {show_text = FALSE}

```

```{r conditional_block2, eval=show_text, echo=FALSE}
#this text prints if there is no available tier III data 
# cat("There is currently no available modeled sediment carbon stock estimates for mangrove or marsh habitats in", chosen_geog)
```

```{r Tier3 Overlap, echo=FALSE}

#creating text and table for tier III overlap and gtlt tier I and II

# tier_overlap <- selectCountry(tier3data) %>% 
#   select(-c(stock_MgHa_mean, stock_MgHa_lowerCI, stock_MgHa_upperCI, tier))
# 
# kable(tier_overlap)
```
## Visualizations 

### 1. Data Metrics 
```{r data metrics, echo=FALSE, warning=FALSE}
map_input %>% 
     dplyr::count(carbon_pool, habitat, territory) %>% 
     plot_ly(x = ~habitat, y = ~n, type = "bar",
             color = ~carbon_pool) %>% 
     layout(xaxis = list(title = "Habitat Type"),
            yaxis = list(title = "Number of Samples"))

```

### 2. Emissions Factors 
```{r emissions factors, echo=FALSE, warning=FALSE}
allstocks %>% 
  select(-c(contains("gtlt"), contains("overlaps"), "TierIorII", "text_position")) %>%
  select(habitat, contains("stock"), contains("Tier")) %>% 
  select(-contains("Total")) %>% 
  pivot_longer(cols = -habitat, names_to = "tier", values_to = "stock") %>% 
  separate(tier, into = c("carbon_pool", "tier", "stat"), sep = "_") %>% 
  pivot_wider(id_cols = c("habitat", "carbon_pool", "tier"), names_from = stat, values_from = stock) %>% 
  
  ggplot2::ggplot(aes(mean, habitat, col = tier)) +
  # geom_boxplot(aes(stock_MgHa, habitat, col = `carbon pool`)) +
  geom_errorbar(aes(xmin = lowerCI, xmax  = upperCI), width = 0.1) +
  geom_point(size = 2, shape = 21, fill="white") +
  theme_bw() +
  facet_wrap(~`carbon_pool`, 
             # scales = "free", 
             dir = "v")
  

```

### 3. Activity Data 
```{r, echo=FALSE, warning=FALSE, results='hide', fig.keep='all'}
    
allstocks %>% 
      plot_ly(x = ~habitat, y = ~area_ha, type = "bar",
              # color = ~habitat, colors = "Set1",
              error_y = ~list(array = area_ha_upperCI - area_ha, 
                              arrayminus = area_ha - area_ha_lowerCI, color = "black")) %>% 
      layout(xaxis = list(title = "Habitat Type"),
             yaxis = list(title = "Area (ha)"))

```

```{r, echo=FALSE, results='asis'}

#cat(
# "## Materials and Methodology 
# 
# This page details the data sources and methodology used to develop and estimate habitat areas and sediment carbon stocks, including data sources and term definitions. 
# 
# ### Disclaimer
# 
# Country and territory names used in this map and dashboard have been aligned with current ESRI guidelines.The boundaries and names shown and the designations used on this map do not imply the expression of any opinion whatsoever on the part of the Coastal Carbon Network or the Smithsonian Institution concerning the legal status of any country, territory, city or area or of its authorities, or concerning the delimitation of its frontiers or boundaries. 
# 
# ### Data Sources 
# 
# Multiple data sources were compiled in order to synthesize data for this tool. Worthington et al 2021 was used to estimate salt marsh area data, Bunting et al 2018 was used to estimate mangrove area data, and McKenzie et al 2021 was used to estimate seagrass area data, when available. Soil and vegetation data available in the Coastal Carbon Atlas, hosted by the Coastal Carbon Network, was used to estimate mean country-level sediment carbon stocks. This data was filtered to exclude cores shorter than 1 meter, in order to align with IPCC global carbon stock value standards. 
# 
# Modeled carbon stock values, estimated Tier III sediment carbon stocks, were estimated from Maxwell et al 2021 and Sanderman et al 2018 for all applicable territories. These estimated values were derived from remote sensing data present in these two syntheses, Maxwell et al 2021 for mangrove habitats, and Sanderman et al 2018 for tidal salt marsh habitats. 
# 
# When neither Tier II or Tier III values were available for a habitat in a geographic territory, the global IPCC sediment carbon stock value for that habitat is applied. 
# 
# ### Mapping 
# 
# The interactive map featured in this tool was created using the leaflet package in R. OSM (default) and CartoDB provider tiles were used to create a global background layer. 
# 
# EEZ data source and Countries data source.
# Spatial interpolation of unmonitored territories/watersheds. 
# 
# ### Estimating Sediment Carbon Stocks
# 
# To estimate territory-level sediment carbon stocks, data from the Coastal Carbon Atlas and Data Library was used. 
# 
# ### Definitions: Total Stocks 
# 
# Habitat Area (ha) : In hectares, area of specified habitat in selected territory.
# 
# Mean Stock (Mg/ha) : Mean soil carbon stock value in milligrams per hectare
# 
# Mean Stock Upper CI (Mg/ha) : Upper confidence interval for mean soil carbon stock value
# 
# Mean Stock Lower CI (Mg/ha): Lower confidence interval for mean soil carbon stock value 
# 
# CO2eq (TgC): (column name will be changed?) 
# 
# Reporting Tier: Level of reporting resolution for carbon stocks. Tiers listed include IPCC global values, country-level stock calculations, and modeled values. Tier I stocks correspond to IPCC global values, Tier II stocks correspond to country-level stock calculations, Tier III stocks correspond to modeled values. The reporting tier listed is the highest available resolution for a given territory and habitat. 
# 
# Reporting Insight: Specifies whether country-specific or modeled carbon stock averages overlap with IPCC global values")
```


## References
