---
title: " `r chosen_country` Country Insights"
output:
  pdf_document:
date: "`r format(Sys.time(), '%d %b %Y')`"
bibliography: all_citations_no_country.bib
engine: knitr
---

```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(knitr)

#specify chosen country 
#chosen_country <- input$chosen_country

#pull in curated app data, filtered to country
all_stocks <- read_csv("all_stocks_table.csv") %>% filter(country == chosen_country)
app_data <- readRDS("app_data.rds") 
area <- app_data$landuse %>% filter(country == chosen_country)
tier1data <- app_data$tier1data %>% filter(country == chosen_country)
tier2data <- app_data$tier2data %>% filter(country == chosen_country)
tier3data <- app_data$tier3data %>% filter(country == chosen_country)
emissionsfactors <- app_data$emissionsfactors %>% filter(country == chosen_country)
totalstocks <- app_data$totalstocks %>% filter(country == chosen_country)
#map_input <- app_data$map_input %>% filter(country == chosen_country)
map_cores <- app_data$map_cores %>% filter(country == chosen_country)
citations <- app_data$citations

```


```{r country citations, echo=FALSE}

#create list of citations by country
country_citations <- citations %>% filter(country == chosen_country)

citation.list <- (unique(country_citations$bibliography_id))

```



```{r habitat areas, echo=FALSE} 
#Area totals for individual habitats 
#mangrove values 
area_ha_mangrove <- area %>% filter(habitat == "mangrove") %>% select(area_ha) 
area_low_mangrove <- area %>% filter(habitat == "mangrove") %>% select(hectare_LowerCI)
area_high_mangrove <- area %>% filter(habitat == "marsh") %>% select(hectare_UpperCI)

#marsh values
area_ha_marsh <- area %>% filter(habitat == "marsh") %>% select(area_ha)
area_low_marsh <- area %>%  filter(habitat == "marsh") %>% select(hectare_LowerCI)
area_high_marsh <- area %>% filter(habitat == "marsh") %>% select(hectare_UpperCI)

#seagrass values 
area_ha_seagrass <- area %>% filter(habitat == "seagrass") %>% select(area_ha)
area_low_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_LowerCI)
area_high_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_UpperCI)
```

```{r total stocks, echo=FALSE}
#total stocks
stocks_country <- totalstocks %>% 
  mutate(total_stocks = round(total_stocks, 2),
         total_stocks_lower = round(total_stocks_lower, 2),
         total_stocks_upper = round(total_stocks_upper, 2),
         total_stocks_se = round(total_stocks_se, 2))

total_stock <- stocks_country %>% filter(habitat == "total") %>% select(total_stocks)
totalstock_country_lower <- stocks_country %>% filter(habitat == "total") %>% select(total_stocks_lower)
totalstock_country_upper <- stocks_country %>% filter(habitat == "total") %>% select(total_stocks_upper)

#total stocks per habitat per country
totalstocks_mangrove <- stocks_country %>% filter(habitat == "mangrove") %>% select(total_stocks)
totalstocks_mangrove_upper <- stocks_country %>% filter(habitat == "mangrove") %>% select(total_stocks_upper)
totalstocks_mangrove_lower <- stocks_country%>% filter(habitat == "mangrove") %>% select(total_stocks_lower)
  
totalstocks_marsh <- stocks_country %>% filter(habitat == "marsh") %>% select(total_stocks)
totalstocks_marsh_upper <- stocks_country %>% filter(habitat == "marsh") %>% select(total_stocks_upper)
totalstocks_marsh_lower <- stocks_country %>% filter(habitat == "marsh") %>% select(total_stocks_lower)
  
totalstocks_seagrass <- stocks_country %>% filter(habitat == "seagrass") %>% select(total_stocks)
totalstocks_seagrass_upper <- stocks_country %>% filter(habitat == "seagrass") %>% select(total_stocks_upper)
totalstocks_seagrass_lower <- stocks_country%>% filter(habitat == "seagrass") %>% select(total_stocks_lower)

```

## Congratulations! This country has available data.

This page includes country-specific insights and more detailed analysis, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass ecosystems. This report details information for the selected country, **`r chosen_country`**.

Please explore the rest of the dashboard for more exciting visualizations, map features, and data. 

Resources referenced to calculate estimates for **`r chosen_country`** are listed below under 'References' at the bottom of this document. 


## Total Carbon Stock Estimates 

Total Carbon stock estimates were calculated for each country and habitat
At this time total Carbon stock estimates do not include seagrass

We estimate that **`r chosen_country`** contains between `r round(totalstock_country_lower, 2)` to 
`r totalstock_country_upper` metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r total_stock` metric tonnes C. 

```{r table1, echo=FALSE}
totals <- totalstocks %>% filter(habitat == "total")
kable(totals)
```

This total estimate includes total mangrove carbon stocks, from `r totalstocks_mangrove_lower` to `r totalstocks_mangrove_upper` metric tonnes of soil C to a depth of 1 m, with
a mean estimate of `r totalstocks_mangrove`

```{r table2, echo=FALSE}
mangrove_totals <- totalstocks %>% filter(habitat == "mangrove")
kable(mangrove_totals)
```

This total estimate also includes total tidal marsh carbon stocks, ranging from `r totalstocks_marsh_lower` to `r totalstocks_marsh_upper`metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r totalstocks_marsh`

```{r table3, echo=FALSE}
marsh_totals <- totalstocks %>% filter(habitat == "marsh")
kable(marsh_totals)
```

Seagrass carbon stocks were not included in the total value due to lack of a global, transparent, and independently assessed seagrass habitat map, however, best available areas and stocks for **`r chosen_country`** are explored in the following 'Wetland Areas and Activities' section. 


## Wetland Areas and Activities

We estimate mangrove area in **`r chosen_country`** to be `r area_low_mangrove` to `r area_high_mangrove` hectares, with a mean estimate of `r area_ha_mangrove` hectares according to Global Mangrove Watch @Bunting_et_al_2018.

We estimate tidal marsh area in **`r chosen_country`** to be `r area_low_marsh` to `r area_high_marsh` hectares, with a mean estimate of hectares according to @Worthington_et_al_2024.

We estimate seagrass area to be **`r chosen_country`** to be a mean of `r area_ha_seagrass` hectares, according to @McKenzie_et_al_2020, aggregating data from multiple sources. 

@McKenzie_et_al_2020 classifies seagrass area estimates as either high or medium to low confidence. seagrass_area_high_confidence % of the estimated seagrass area of **`r chosen_country`** is considered high to medium confidence, while seagrass_area_low_confidence % of the estimated seagrass area is categorized as low confidence. 


## Calculated Stocks and Emissions Factors 

```{r N_cores, echo = FALSE}

Ncores <- map_cores %>% filter(carbon_pool == "soil") %>% count() 
Nsites <- map_cores %>% filter(carbon_pool == "soil") %>% select(site_id) %>% distinct() %>% count() 
Nhabitats <- map_cores %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct() %>% count() 

```

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_country`**. 

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas. 

Data from **`r chosen_country`** includes `r Ncores` soil profiles from `r Nsites`watersheds. This data comes from `r Nhabitats` different habitat types. 

If there is not yet any country specific information in the Coastal Carbon Atlas, we instead applied IPCC Tier I estimate. IPCC Tier I estimates for mangrove, marsh, and seagrass ecosystems are listed below. **SOURCE**

The table in this section also details whether the calculated Tier II value is significantly different from the estimated Tier I values. This is observed in the "Overlap" column.  

```{r IPCC, echo =FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- as_tibble(data.frame(Habitat = c("mangrove", "marsh", "seagrass"),
                                      Mean = c(386, 255, 108),
                                      `Lower_CI` = c(351,254,84),
                                      `Upper_CI` = c(424,297,139)))

kable(IPCC_tier_I, caption = "IPCC Tier I Value Estimates")
```


```{r conditional_print, echo = FALSE}
#create table of available tier estimates and any overlap 
country_tier <- all_stocks %>% filter(country == chosen_country)

tier_mangrove <- country_tier %>% filter(habitat == "mangrove") %>% select(TierIorII)
tier_marsh <- country_tier %>% filter(habitat == "marsh") %>% select(TierIorII)
tier_seagrass <- country_tier %>% filter(habitat == "seagrass") %>% select(TierIorII)

tiertable <- country_tier %>% select(country, territory, habitat, TierIorII, tier_II_overlaps_TierI) %>% 
  rename(Country = country, 
         Territory = territory,
         Habitat = habitat,
         Tier = TierIorII,
         Overlap = tier_II_overlaps_TierI)

kable(tiertable, caption = "Availiability of Tier I and Tier II Data")

```

### Tier I Carbon Stocks 

This table includes Tier I Carbon Stocks included for **`r chosen_country`**. 

```{r tier1, echo=FALSE}
tier1stocks <- tier1data %>% filter(country == chosen_country)
kable(tier1stocks)
```


### Tier II Carbon Stocks 

This table includes Tier II Carbon Stock estimates for **`r chosen_country`**. Estimates in this table were derived from data queried from the Coastal Carbon Atlas. SOURCE

```{r tier2, echo=FALSE}
library(knitr)
tier2stocks <- tier2data %>% filter(country == chosen_country)
kable(tier2stocks)
```


### Tier III Carbon Stocks 

Tier III carbon stocks were estimated, when available, from remote sensing data from Maxwell et al 2021 and Sanderman et al 2018. The table below details whether estimated values are available for **`r chosen_country`**, and any overlap with associated Tier I or Tier II values. 

If there are no Tier III estimates associated with the selected country, please refer to Tier I and Tier II tables.

```{r tier3, echo=FALSE}
#tier3stocks <- tier3data %>% filter(country == country_selected)
#if no 

if(chosen_country %in% tier3data$country){
  
      tierIII <- tier3data %>% 
        filter(country == chosen_country)
        kable(tierIII)
      
    } else {
      print("There are currently no Tier III remote sensing estimates for this country. Please refer to Tier I and Tier II carbon stock estimates.")
    }


# if(chosen_country %in% tier3data$country){
#   
#   #want to pull greater than/less than and overlap values here 
#  
#     } else {
#       print("There are currently no Tier III remote sensing estimates for this country. Please refer to Tier I and Tier II carbon stock estimates.")
#     }

```



## References 


