---
title: "Country Insights: Country"
output:
  pdf_document:
    toc: true
engine: knitr
---

```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)

### This script writes the country insights pdf for the case of countries with AVAILABLE Tier II DATA 

#specify chosen country 
#chosen_country <- input$chosen_country

#for script testing 
chosen_country <- "Australia"



#function to select only observations with chosen country 
selectCountry <- function(df){
  df %>% filter(country == chosen_country)
}

#pull in curated app data, filtered to country
stocks <- read_csv("data/all_stocks_table.csv") 
all_stocks <- selectCountry(stocks)
app_data <- readRDS("data/app_data.rds") 
area <- app_data$landuse %>% selectCountry()
tier1data <- app_data$tier1data 
tier2data <- app_data$tier2data
tier3data <- app_data$tier3data
emissionsfactors <- app_data$emissionsfactors %>% selectCountry()
totalstocks <- app_data$totalstocks %>% selectCountry()
#map_input <- app_data$map_input %>% filter(country == chosen_country)
map_cores <- app_data$map_cores %>% selectCountry()
citations <- app_data$citations

rm(app_data) #keeps the rds out of the environment 

```

```{r country citations, echo=FALSE}

#create list of citations by country
country_citations <- selectCountry(citations)
citation.list <- (unique(country_citations$bibliography_id))

```

```{r habitat areas, echo=FALSE}
#Area totals for individual habitats 
#mangrove values 
area_ha_mangrove <- area %>% filter(habitat == "mangrove") %>% select(area_ha) 
area_low_mangrove <- area %>% filter(habitat == "mangrove") %>% select(hectare_LowerCI)
area_high_mangrove <- area %>% filter(habitat == "marsh") %>% select(hectare_UpperCI)

#marsh values
area_ha_marsh <- area %>% filter(habitat == "marsh") %>% select(area_ha)
area_low_marsh <- area %>%  filter(habitat == "marsh") %>% select(hectare_LowerCI)
area_high_marsh <- area %>% filter(habitat == "marsh") %>% select(hectare_UpperCI)

#seagrass values 
area_ha_seagrass <- area %>% filter(habitat == "seagrass") %>% select(area_ha)
area_low_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_LowerCI)
area_high_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_UpperCI)
```

```{r total stocks, echo=FALSE}
#total stocks

totalstocks <- totalstocks %>% 
  mutate(total_stocks = signif(total_stocks, 6),
         total_stocks_lower = signif(total_stocks_lower, 6),
         total_stocks_upper = signif(total_stocks_upper, 6),
         total_stocks_se = signif(total_stocks_se, 6))


total_stock <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks)
totalstock_country_lower <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks_lower)
totalstock_country_upper <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks_upper)

#total stocks per habitat per country
totalstocks_mangrove <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks)
totalstocks_mangrove_upper <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks_upper)
totalstocks_mangrove_lower <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks_lower)
  
totalstocks_marsh <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks)
totalstocks_marsh_upper <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks_upper)
totalstocks_marsh_lower <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks_lower)
  
totalstocks_seagrass <- totalstocks %>% filter(habitat == "seagrass") %>% select(total_stocks)
totalstocks_seagrass_upper <- totalstocks %>% filter(habitat == "seagrass") %>% select(total_stocks_upper)
totalstocks_seagrass_lower <- totalstocks%>% filter(habitat == "seagrass") %>% select(total_stocks_lower)

```

## Welcome to the CCN Inventory Tool Country Insights

Congratulations! This country has available data.

This document includes country-specific insights and more detailed analysis, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass habitats. This report details information for the selected country, **`r chosen_country`**.

Explore the rest of the dashboard for more exciting visualizations, map features and data.

Resources referenced to calculate estimates for **`r chosen_country`** are listed below under 'References' at the end of this document.

If you have any questions or data you would like to add to the Coastal Carbon Network, please reach out to us at [CoastalCarbon\@si.edu](mailto:CoastalCarbon@si.edu){.email}.

## Total Carbon Stock Estimates

Total Carbon stock estimates were calculated for each country in total, and for mangrove, tidal marsh, and seagrass habitats. At this time total Carbon stock estimates do not include seagrass estimates due to MORE HERE 

We estimate that **`r chosen_country`** contains a mean estimate of `r signif(total_stock, 6)` metric tonnes soil Carbon.

We estimate that **`r chosen_country`** contains between `r signif(totalstock_country_lower, 6)` to `r signif(totalstock_country_upper, 6)` metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r signif(total_stock, 6)` metric tonnes C.

```{r table1 total, echo=FALSE}

totals <- totalstocks %>% filter(habitat == "total") %>% 
  mutate(Country = country,
         Territory = territory,
         Habitat = habitat,
         `Total Stock` = signif(total_stocks, 6),
         `Lower CI` = signif(total_stocks_lower, 6),
         `Upper CI` = signif(total_stocks_upper, 6),
         `Standard Error` = total_stocks_se) %>% 
  janitor::remove_empty(which = c("cols")) %>% 
  select(-c(country, territory, habitat, total_stocks, total_stocks_lower, total_stocks_upper,total_stocks_se)) %>% 
  select(Country, Territory, Habitat, `Total Stock`, everything())


kable(totals)
```

#if mangrove data

This total estimate includes total mangrove soil carbon stocks, from `r signif(totalstocks_mangrove_lower, 6)` to `r signif(totalstocks_mangrove_upper, 6)` metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r signif(totalstocks_mangrove, 6)`

```{r table2 mangrove, echo=FALSE}
mangrove_totals <- totalstocks %>% filter(habitat == "mangrove") %>% 
  rename(`Lower CI` = total_stocks_lower,
         `Upper CI` = total_stocks_upper,
         `Standard Error` = total_stocks_se) %>% 
   mutate(Country = country,
          Territory = territory,
          Habitat = habitat,
          `Total Stock` = signif(total_stocks, 6),
          `Lower CI` = signif(`Lower CI`, 6),
          `Upper CI` = signif(`Upper CI`, 6)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(-c(country, territory, habitat, total_stocks)) %>% 
  select(Country, Territory, Habitat, `Total Stock`, everything())

kable(mangrove_totals)
```

#if no mangrove data

This total estimate includes total mangrove soil carbon stocks, with a mean estimate of `r signif(totalstocks_mangrove, 6)`.

This total estimate also includes total tidal marsh carbon stocks, ranging from `r signif(totalstocks_marsh_lower, 6)` to `r signif(totalstocks_marsh_upper, 6)` metric tonnes of soil C to a depth of 1 m, with a mean estimate of `r signif(totalstocks_marsh, 6)`

```{r table3 marsh, echo=FALSE}

marsh_totals <- totalstocks %>% filter(habitat == "marsh") %>% 
  rename(`Lower CI` = total_stocks_lower,
         `Upper CI` = total_stocks_upper,
         `Standard Error` = total_stocks_se) %>% 
   mutate(Country = country,
          Territory = territory,
          Habitat = habitat,
          `Total Stock` = signif(total_stocks, 6),
          `Lower CI` = signif(`Lower CI`, 6),
          `Upper CI` = signif(`Upper CI`, 6)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(-c(country, territory, habitat, total_stocks)) %>% 
  select(Country, Territory, Habitat, `Total Stock`, everything())

kable(marsh_totals)
```

We estimate seagrass soil carbon stocks for **`r chosen_country`** to have a mean estimate of `r totalstocks_seagrass`

```{r table4 seagrass, echo=FALSE}

seagrass_totals <- totalstocks %>% filter(habitat == "seagrass") %>% 
  rename(`Lower CI` = total_stocks_lower,
         `Upper CI` = total_stocks_upper,
         `Standard Error` = total_stocks_se) %>% 
   mutate(Country = country,
          Territory = territory,
          Habitat = habitat,
          `Total Stock` = signif(total_stocks, 6),
          `Lower CI` = signif(`Lower CI`, 6),
          `Upper CI` = signif(`Upper CI`, 6)) %>% 
  janitor::remove_empty(which = c("cols")) %>% #removes any empty columns if country is missing data 
  select(-c(country, territory, habitat, total_stocks))

kable(seagrass_totals)
```


## Wetland Areas and Activities

We estimate mangrove area in **`r chosen_country`** to be `r signif(area_low_mangrove, 4)` to `r signif(area_high_mangrove, 4)` hectares, with a mean estimate of `r signif(area_ha_mangrove, 4)` hectares according to Global Mangrove Watch @Bunting_et_al_2018.

We estimate tidal marsh area in **`r chosen_country`** to be `r signif(area_low_marsh, 4)` to `r signif(area_high_marsh, 4)` hectares, with a mean estimate of hectares according to @Worthington_et_al_2024.

We estimate seagrass area to be **`r chosen_country`** to be a mean of `r signif(area_ha_seagrass, 4)` hectares, according to @McKenzie_et_al_2020, which aggregates global seagrass data from a number of sources. 

@McKenzie_et_al_2020 classifies seagrass area estimates as either high or medium-to-low confidence. seagrass_area_high_confidence % of the estimated seagrass area of **`r chosen_country`** is considered high to medium confidence, while seagrass_area_low_confidence % of the estimated seagrass area is categorized as low confidence.

## Calculated Stocks and Emissions Factors

```{r N_cores, echo = FALSE}
#cores, sites, habitats for soil cores included in tier II calcs 
cores <- selectCountry(map_cores) %>% filter(carbon_pool == "soil") %>% count() 
sites <- selectCountry(map_cores) %>% filter(carbon_pool == "soil") %>% select(site_id) %>% distinct() %>% count() 
habitats <- selectCountry(map_cores) %>% filter(carbon_pool == "soil") %>% select(habitat) %>% distinct() %>% count() 

```

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_country`**.

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas.

Data from **`r chosen_country`** includes `r cores` soil profiles from `r sites` watersheds. This data comes from `r habitats` different habitat types.

In the case that there is not yet any country specific information in the Coastal Carbon Atlas, we instead applied IPCC Tier I estimate. IPCC Tier I estimates for mangrove, marsh, and seagrass ecosystems are listed below. **SOURCE**

The table in this section also details whether the calculated Tier II value is significantly different from the estimated Tier I values. This is observed in the "Overlap" column.

```{r IPCC, echo =FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- as_tibble(data.frame(Habitat = c("mangrove", "marsh", "seagrass"),
                                      Mean = c(386, 255, 108),
                                      `Lower_CI` = c(351,254,84),
                                      `Upper_CI` = c(424,297,139)))

kable(IPCC_tier_I, caption = "IPCC Tier I Value Estimates")
```

```{r tier availability, echo = FALSE}
#create table of available tier estimates and any overlap 

tier_mangrove <- all_stocks %>% filter(habitat == "mangrove") %>% select(TierIorII)
tier_marsh <- all_stocks %>% filter(habitat == "marsh") %>% select(TierIorII)
tier_seagrass <- all_stocks %>% filter(habitat == "seagrass") %>% select(TierIorII)

tiertable <- all_stocks %>% select(country, territory, habitat, TierIorII, tier_II_overlaps_TierI) %>% 
  rename(Country = country, 
         Territory = territory,
         Habitat = habitat,
         Tier = TierIorII,
         Overlap = tier_II_overlaps_TierI) %>% 
  mutate(Overlap = if_else(is.na(Overlap), 
        "No Tier II data available at this time", Overlap))

kable(tiertable, caption = "Availiability of Tier I and Tier II Data")

```

### Tier I Carbon Stocks

This section includes Tier I Carbon Stocks included for **`r chosen_country`**. In the case that all habitats; mangrove, tidal marsh, and seagrass have availiable Tier II estimates, please refer to Table 5: IPCC Tier I Value Estimates for applicable Tier I values. 

```{r setuptier1, echo=FALSE, warning=FALSE, message=FALSE}
  if (chosen_country %in% tier1data$country)
     {show_text = TRUE
     } else {show_text = FALSE}
```

```{r tier1eval, eval=show_text, echo=FALSE}
 tier1stocks <- selectCountry(tier1data) %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = stock_MgHa_mean,
        `Lower CI` = stock_MgHa_lowerCI,
        `Upper CI` = stock_MgHa_upperCI) %>% 
  janitor::remove_empty(which = "cols") %>%  #just in case some countries have missing variables 
  select(-c(tier, carbon_pool))


kable(tier1stocks, caption = "Tier I Carbon Stock Estimates")
```

```{r setuptier1_false, echo=FALSE, warning=FALSE, message=FALSE}
  if (!chosen_country %in% tier1data$country)
     {show_text = TRUE
     } else {show_text = FALSE}
```

```{r tier1eval2, eval=show_text, echo=FALSE, results='asis'}
cat("Good news! The selected country has Tier II data for each available habitat. Please refer to Table 5: IPCC Tier I Value Estimates for applicable Tier I estimates and Table 7: Tier II Carbon Stock Estimates.")
```


### Tier II Carbon Stocks

This table includes Tier II Carbon Stock estimates for **`r chosen_country`**. Estimates in this table were derived from data queried from the Coastal Carbon Atlas. SOURCE

A selected country may have available Tier II values for one or multiple habitats, this is dependent on cores accessible through the Coastal Carbon Atlas. At the time of analysis, referencing Version 1.5.0, **`r chosen_country`** contains a total of `r cores` cores across `r habitats` unique habitats.

```{r tier2, echo=FALSE}
library(knitr)
tier2stocks <- selectCountry(tier2data) %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = stock_MgHa_mean,
        `Lower CI` = stock_MgHa_lowerCI,
        `Upper CI` = stock_MgHa_upperCI,
        `Standard Error` = stock_MgHa_se) %>%
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`, 3)) %>% 
  janitor::remove_empty(which = "cols") %>% #removing any habitats without tier II data 
  select(-c(tier, carbon_pool))

kable(tier2stocks, caption = "Tier II Carbon Stock Estimates") 
```

### Tier III Carbon Stocks

Tier III carbon stocks were estimated, when available, from remote sensing data from Maxwell et al 2021 and Sanderman et al 2018. The table below details whether estimated values are available for **`r chosen_country`**, and any overlap with associated Tier I or Tier II values.

If there are no Tier III estimates associated with the selected country and specific habitat of interest, please refer to above Tier I and Tier II tables.

```{r tier 3setup1, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is tier3 data for at least one habitat 
  if (chosen_country %in% tier3data$country)
     {show_text = TRUE
     } else {show_text = FALSE}

```

```{r tier overlap, echo=FALSE}
#tier 3 overlap with 1 and 2
tier_overlap <- selectCountry(tier3data) %>% 
  select(-c(stock_MgHa_mean, stock_MgHa_lowerCI, stock_MgHa_upperCI, tier))

#create variables to use in inline code
tierIII_gtlt_tier_II <- tier_overlap %>% select(tierIII_gtlt_tier_II)
tierII_overlaps_tierIII <- tier_overlap %>% select(tierII_overlaps_tierIII)  
tierIII_gtlt_tier_I <- tier_overlap %>% select(tierIII_gtlt_tier_I)
tierIII_overlaps_tierI <- tier_overlap %>% select(tierIII_overlaps_tierI)
```

```{r conditional_block1, eval=show_text, echo=FALSE, results='asis'}

tier3stocks <- selectCountry(tier3data) %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = stock_MgHa_mean,
        `Lower CI` = stock_MgHa_lowerCI,
        `Upper CI` = stock_MgHa_upperCI) %>% 
  mutate(`Mean Stock (MgHa)` = signif(`Mean Stock (MgHa)`, 3),
         `Lower CI` = signif(`Lower CI`, 3),
         `Upper CI` = signif(`Upper CI`,3)) %>% 
  select(-c(tier, tierIII_gtlt_tier_II, tierII_overlaps_tierIII,
            tierIII_gtlt_tier_I, tierIII_overlaps_tierI))

 kable(tier3stocks, caption = "Tier III Carbon Stock Estimates") 

tier_overlap <- selectCountry(tier3data) %>% 
  select(-c(stock_MgHa_mean, stock_MgHa_lowerCI, stock_MgHa_upperCI, tier))

#create variables to use in inline code
text1 <- tier_overlap %>% select(tierIII_gtlt_tier_II) %>% as.character()
text2 <- tier_overlap %>% select(tierII_overlaps_tierIII) %>% as.character()
text3  <- tier_overlap %>% select(tierIII_gtlt_tier_I) %>% as.character()
text4 <- tier_overlap %>% select(tierIII_overlaps_tierI) %>% as.character()
 
cat("In",chosen_country,", Tier III estimated stock MgHa values are", text1, "Tier II values estimated from Coastal Carbon Atlas Data. In this case, the Tier III", text2)

cat(". Additionally, estimated Tier III stock values derived from remote sensing are", text3, "estimates Tier I values.
    In this case, the Tier III", text4)

```


```{r tier3setup2, echo=FALSE, warning=FALSE, message=FALSE}
#this chunk sets up the condition that there is NO tier 3 data 
  if (!chosen_country %in% tier3data$country)
     {show_text = TRUE
     } else {show_text = FALSE}

```

```{r conditional_block2, eval=show_text, echo=FALSE}
#this text prints if there is no available tier III data 
cat("There is currently no available Tier III level data for mangrove or marsh habitats in" chosen_country)
```



```{r Tier3 Overlap, echo=FALSE}

#creating text and table for tier III overlap and gtlt tier I and II

# tier_overlap <- selectCountry(tier3data) %>% 
#   select(-c(stock_MgHa_mean, stock_MgHa_lowerCI, stock_MgHa_upperCI, tier))
# 
# kable(tier_overlap)
```






## References
