---
title: "Country Insights: `r chosen_country`"
output:
  pdf_document:
    toc: true
engine: knitr
---

```{r setup, include=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)

### This script writes the country insights pdf for the case of countries with NO AVAILABLE Tier II DATA 

#function to select only observations with chosen country 
selectCountry <- function(df){
  df %>% filter(country == chosen_country)
}


#for script testing 

#pull in curated app data, filtered to country
stocks <- read_csv("data/all_stocks_table.csv") 
all_stocks <- selectCountry(stocks)
app_data <- readRDS("data/app_data.rds") 
area <- app_data$landuse %>% selectCountry()
tier1data <- app_data$tier1data 
tier2data <- app_data$tier2data
tier3data <- app_data$tier3data
emissionsfactors <- app_data$emissionsfactors %>% selectCountry()
totalstocks <- app_data$totalstocks %>% selectCountry()
#map_input <- app_data$map_input %>% filter(country == chosen_country)
map_cores <- app_data$map_cores %>% selectCountry()
citations <- app_data$citations

rm(app_data) #keeps the rds out of the environment 

```

```{r country citations, echo=FALSE}

#create list of citations by country
country_citations <- selectCountry(citations)
citation.list <- (unique(country_citations$bibliography_id))

```

```{r habitat areas, echo=FALSE}
#Area totals for individual habitats 
#mangrove values 
area_ha_mangrove <- area %>% filter(habitat == "mangrove") %>% select(area_ha) 
area_low_mangrove <- area %>% filter(habitat == "mangrove") %>% select(hectare_LowerCI)
area_high_mangrove <- area %>% filter(habitat == "mangrove") %>% select(hectare_UpperCI)

#marsh values
area_ha_marsh <- area %>% filter(habitat == "marsh") %>% select(area_ha)
area_low_marsh <- area %>%  filter(habitat == "marsh") %>% select(hectare_LowerCI)
area_high_marsh <- area %>% filter(habitat == "marsh") %>% select(hectare_UpperCI)

#seagrass values 
area_ha_seagrass <- area %>% filter(habitat == "seagrass") %>% select(area_ha)
area_low_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_LowerCI)
area_high_seagrass <- area %>% filter(habitat == "seagrass") %>% select(hectare_UpperCI)
```

```{r total stocks, echo=FALSE}
#total stocks
total_stock <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks)
totalstock_country_lower <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks_lower)
totalstock_country_upper <- totalstocks %>% filter(habitat == "total") %>% select(total_stocks_upper)

#total stocks per habitat per country
totalstocks_mangrove <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks)
totalstocks_mangrove_upper <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks_upper)
totalstocks_mangrove_lower <- totalstocks %>% filter(habitat == "mangrove") %>% select(total_stocks_lower)
  
totalstocks_marsh <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks)
totalstocks_marsh_upper <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks_upper)
totalstocks_marsh_lower <- totalstocks %>% filter(habitat == "marsh") %>% select(total_stocks_lower)
  
totalstocks_seagrass <- totalstocks %>% filter(habitat == "seagrass") %>% select(total_stocks)
totalstocks_seagrass_upper <- totalstocks %>% filter(habitat == "seagrass") %>% select(total_stocks_upper)
totalstocks_seagrass_lower <- totalstocks%>% filter(habitat == "seagrass") %>% select(total_stocks_lower)

```

## Welcome!

You have reached the CCN Inventory Tool Country Insights Report.

Congratulations! This country has available data. #Fix this description?

This is a downloadable document which includes country-specific insights and more detailed analysis, including carbon stocks, emissions factors, and ecosystem wetland area for mangrove, marsh, and seagrass habitats. This report details information for the selected country, **`r chosen_country`**.

Explore the rest of the dashboard for more exciting visualizations, map features, and data.

Resources referenced to calculate Carbon stock estimates for **`r chosen_country`** are listed below 'References' at the end of this document.

If you have any questions or data you would like to add to the Coastal Carbon Network, please reach out to us at [CoastalCarbon\@si.edu](mailto:CoastalCarbon@si.edu){.email}.

## Total Carbon Stock Estimates

Total Carbon stock estimates were calculated for each country in total, and for mangrove, tidal marsh, and seagrass habitats. At this time total Carbon stock estimates for each country do not include seagrass estimates due to the lack of a global, transparent, and independently assessed seagrass habitat map. However, best available areas and stocks for **`r chosen_country`** are explored in the following 'Wetland Areas and Activities' section.

### Country Level Total Stock Estimates

We estimate that **`r chosen_country`** contains between `r signif(totalstock_country_upper, 6)` to `r signif(totalstock_country_lower, 6)` metric tonnes of soil Carbon to a depth of 1 meter, with a mean estimate of `r signif(total_stock, 6)` metric tonnes soil Carbon.

```{r table1, echo=FALSE}

totals <- totalstocks %>% filter(habitat == "total") %>% 
  mutate(Country = country,
         Territory = territory,
         Habitat = habitat,
         `Total Stock` = signif(total_stocks, 6),
         `Upper CI` = signif(total_stocks_lower, 6),#for some reason the labels are switched
         `Lower CI` = signif(total_stocks_upper, 6),
         `Standard Error` = signif(total_stocks_se, 5)) %>% 
  select(-c(country, territory, habitat, total_stocks, total_stocks_lower, total_stocks_upper,total_stocks_se))


kable(totals, caption = "Total Country Level Stocks")
```

### Mangrove Estimates

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of `r signif(totalstocks_mangrove, 6)` metric tonnes soil Carbon in mangrove habitats.

### Tidal Marsh Estimates

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of `r signif(totalstocks_marsh, 6)` metric tonnes soil Carbon in tidal marsh habitats.

### Seagrass Estimates 

This total estimate includes country-level mangrove carbon stock estimates, with a mean estimate of `r signif(totalstocks_seagrass, 6)` metric tonnes soil Carbon in seagrass habitats.


## Wetland Areas and Activities

### Mangrove Area

```{r mangrovearea, echo=FALSE, warning=FALSE, message=FALSE}

area_mangrove <- as.numeric(area_ha_mangrove) %>% signif(4)
area_low_mangrove <- as.numeric(area_low_mangrove) %>% signif(4)
area_high_mangrove <- as.numeric(area_high_mangrove) %>% signif(4)

#evaluating the case of available mangrove area 
  if (area_mangrove > 0)
     {show_text = TRUE
  } else {show_text = FALSE}

  if (area_mangrove <= 0|is.na(area_mangrove))
     {show_text2 = TRUE
  } else {show_text2 = FALSE}

```

```{r mangroveeval, eval=show_text, echo=FALSE, results='asis'}

cat("We estimate mangrove area in",chosen_country, "to be", area_low_mangrove, "to", area_high_mangrove,"hectares with a mean estimate of", area_mangrove,"hectares according to Global Mangrove Watch @Bunting_et_al_2018.")
```

```{r mangroveeval2, eval=show_text2, echo=FALSE, results='asis'}
cat("We estimate mangrove area in", chosen_country, "to be approximately 0 hectares according to Global Mangrove Watch @Bunting_et_al_2018")
```


### Tidal Marsh Area

```{r marsharea, echo=FALSE, warning=FALSE, message=FALSE}

  if (area_ha_marsh > 0)
     {show_text = TRUE
  } else {show_text = FALSE}

if (area_ha_marsh <= 0)
     {show_text2 = TRUE
  } else {show_text2 = FALSE}

```

```{r marsheval, eval=show_text, echo=FALSE, results='asis'}

area_marsh <- as.numeric(area_ha_marsh) %>% signif(4)
area_low_marsh <- as.numeric(area_low_marsh) %>% signif(4)
area_high_marsh <- as.numeric(area_high_marsh) %>% signif(4)

cat("We estimate tidal marsh area in",chosen_country, "to be", area_low_marsh, "to", area_high_marsh,"hectares with a mean estimate of", area_marsh,"hectares according to Global Mangrove Watch @Bunting_et_al_2018.") 
```

```{r marsheval2, eval=show_text2, echo=FALSE, results='asis'}
cat("We estimate tidal marsh area in", chosen_country, "to be approximately 0 hectares according to Worthington et al 2024.")
```

### Seagrass Area
```{r seagrassarea, echo=FALSE, warning=FALSE, message=FALSE}

  if (area_ha_seagrass > 0)
     {show_text = TRUE
  } else {show_text = FALSE}

if (area_ha_seagrass <= 0)
     {show_text2 = TRUE
  } else {show_text2 = FALSE}

```

```{r seagrasseval, eval=show_text, echo=FALSE, results='asis'}

area_seagrass <- as.numeric(area_ha_seagrass) %>% signif(4)

cat("We estimate seagrass area in",chosen_country, "to have a mean estimate of", area_seagrass,"hectares according to @McKenzie_et_al_2020, which aggregates data from multiple sources. ")


cat("@McKenzie_et_al_2020 classifies seagrass area estimates as either high or medium to low confidence. seagrass_area_high_confidence % of the estimated seagrass area of",chosen_country, "is considered high to medium confidence, while seagrass_area_low_confidence % of the estimated seagrass area is categorized as low confidence.")
```

```{r seagrasseval2, eval=show_text2, echo=FALSE, results='asis'}
cat("We estimate seagrass area in", chosen_country, "to be approximately 0 hectares according to McKenzie et al 2020.")
```

## Calculated Stocks and Emissions Factors

This section of the report details whether data is available to estimate Tier I, Tier II, or Tier III value estimates for tidal marsh, mangrove, and seagrass ecosystems in **`r chosen_country`**.

If data for the selected country is available in the Coastal Carbon Atlas, we have applied a Tier II emission factor based on a simple average of country specific data queried from the Atlas. SOURCE

Country-specific soil cores from **`r chosen_country`** have not yet been included in the Coastal Carbon Atlas. In this case, we have instead applied the IPCC Tier I estimated value. IPCC Tier I estimates for mangrove, marsh, and seagrass habitats are listed below. **SOURCE**

```{r IPCC, echo=FALSE}
#include table of IPCC Tier I values 
IPCC_tier_I <- tibble(Habitat = c("mangrove", "marsh", "seagrass"),
                      Mean = c(386, 255, 108),
                      `Lower CI` = c(351,254,84),
                      `Upper CI` = c(424,297,139))

kable(IPCC_tier_I, caption = "IPCC Tier I Value Estimates")
```


### Tier I Carbon Stocks

This table includes Tier I Soil Carbon Stocks included for **`r chosen_country`**.

```{r tier1, echo=FALSE}
tier1stocks <- selectCountry(tier1data) %>% 
  rename(Country = country,
         Territory = territory,
         Habitat = habitat,
        `Mean Stock (MgHa)` = stock_MgHa_mean,
        `Lower CI` = stock_MgHa_lowerCI,
        `Upper CI` = stock_MgHa_upperCI) %>% 
  select(-c(tier, carbon_pool))


kable(tier1stocks)
```

### Tier II Carbon Stocks

Tier II carbon stocks were estimated, when available, from soil core data currently housed in the Coastal Carbon Atlas. 

Tier II Data Availability for **`r chosen_country`**:
The selected country, **`r chosen_country`** does not currently have available Tier II carbon stock estimates at a depth of 1 meter. 


### Tier III Carbon Stocks

Tier III carbon stocks were estimated, when available, from remote sensing data from Maxwell et al 2021 for mangrove habitats, and Sanderman et al 2018 for tidal marsh habitats. The following details whether estimated values are available for **`r chosen_country`**, and any overlap with associated Tier I or Tier II values.

If there are no Tier III estimates associated with the selected country, please refer to Tier I and/or Tier II tables.

Tier III Data Availability for **`r chosen_country`**: 
There are currently no Tier III values for **`r chosen_country`** in the CCN Inventory Dashboard. Please refer to Tier I values.


## References
